### 2005-04-04 paper/metamonads.tex
continue with write up of section 4. add some comments in later sections too.

\documentclass{llncs}

% Reduce display spacing
\divide\abovedisplayskip 2
\divide\belowdisplayskip 2
\divide\abovedisplayshortskip 2
\divide\belowdisplayshortskip 2

\usepackage{hyphenat}
\usepackage{comment}
\usepackage{amsmath,amssymb}
\usepackage{amstext}
\usepackage{url}

\usepackage{ifpdf}
\ifpdf
    \pdfpageheight=11in
    \pdfpagewidth=8.5in
\fi

%\usepackage{refrange}

\usepackage{fancyvrb}
\DefineShortVerb{\|}
\DefineVerbatimEnvironment{code}{Verbatim}{xleftmargin=\mathindent}
\setlength{\parskip}{0pt}
\newlength{\mathindent}
\setlength{\mathindent}{1em}

% \usepackage[backref,colorlinks,bookmarks=true]{hyperref}

% Reduce list spacing
%% \makeatletter
%% \renewcommand\@@listI{\leftmargin\leftmargini
%% \parsep \z@@
%% \topsep 3\p@@ \@@plus\p@@ \@@minus 2\p@@
%% \itemsep 2\p@@ \@@plus\p@@ \@@minus\p@@}
%% \let\@@listi\@@listI
%% \@@listi
%% \makeatother

\newcommand{\omitnow}[1]{}
\newcommand{\oleg}[1]{{\it [Oleg says: #1]}}
\newcommand{\jacques}[1]{{\it [Jacques says: #1]}}

\begin{document}
\title{Functors, CPS and monads, or how to generate efficient
code from abstract designs}
\author{Jacques Carette\inst{1}
 \and
Oleg Kyselyov\inst{2}
}
\institute{McMaster University,
1280 Main St. West, Hamilton, Ontario Canada L8S 4K1
\and
Monterey, CA 93943}

\maketitle

\begin{abstract}
We use monads and Ocaml's advanced module system to
demonstrate that, in a generative context, it is
possible to totally eliminate the overhead of abstraction.
This lets one use extreme forms of \textit{information hiding}
at no run-time cost.  Furthermore the typed nature of
the generative context provide static guarantees about the
generated code.  The various extensions
(aspects) can be made completely orthogonal and compositional,
even in the presence of name-generation for temporaries and
other bindings and ``interleaving'' of aspects.  We also
show how to encode some domain-specific knowledge so that
``clearly wrong'' compositions can be statically rejected by the
compiler.
\end{abstract}

\section{Introduction}

In high-performance and numeric computing, there is a well-known
issue of balancing between maximal performance and the level of
abstraction at which code is written.  Furthermore, already in
linear algebra, there is a wealth of different concerns that 
\emph{may} need to be addressed.  While it is possible to manually
inline and simplify such code, this is labour-intensive and 
error-prone.  Furthermore the resulting code is frequently unreadable,
and thus unmaintainable.

There are a number of projects that have attempted to address these issues
\cite{Czarnecki,Veldhuizen:1998:ISCOPE,musser94algorithmoriented,BOOST,POOMA,ATLAS}, 
and they have generally succeeded in generating
efficient code, but the cost was frequently the loss of modularity
of the design.  In other words, to allow the various aspects to
weave with each other \emph{efficiently}, information which should
have been kept hidden leaked from module to module.  This seriously
affects maintainability and scalability.  Furthermore, current architectures
demand more and more frequent tweaks which, in general, cannot be done by the
compiler because the tweaking often involves domain knowledge.

Our goal was to implement a completely modular, textbook-style 
modular design \cite{journals/cacm/parnas72a}, yet 
have our code be very efficient.  Gaussian elimination, as it is a  
well-known algorithm, already contains many different aspects 
\cite{carette04}, was chosen as a representative example.

Whether one prefers to code in an imperative, object-oriented, functional, 
or even in a mixed style, all abstraction mechanisms in current
languages have a non-zero run-time cost.  Whether it is procedure,
method or function call, these have a cost; worse still, higher level
and/or higher order abstractions (like Functors or AspectJ's aspects)
frequently have a larger cost still.  Of course, if one is faced with
writing 60 different implementations of the ``same'' algorithm, this
abstraction cost starts to look cheap!  Although bad programmers have
been known to resort to cut\&paste programming, no self-respecting
programmer would be caught doing that\footnote{Sadly, there are still
many programmers who do not see the harm of cut\&paste programming}.
The common solution is to turn to code generation.

While many people still make do with string-based code generation,
this is not much different than writing code in a dynamically typed
interpreted scripting language.  It works, it is convenient, but it
does not scale well, it is a debugging nightmare, and usually suffers
most from Software Aging \cite{parnas_aging}.  As with other tasks in
programming, the answer is to turn to statically typed languages,
which provide strong guarantees about the soundness of the resulting
code.  MetaOCaml \cite{CTHL03,metaocaml-org} in particular stands
out: not only it guarantees well-formedness and the absense of problems 
such as accidental variable capture \cite{HygienicMacros}
\jacques{is this the citation you meant?}
MetaOcaml also
statically ensures that both the code generator and the generated code
are well-typed \cite{TahaSheard97,TahaThesis}.  While Template
programming in C++ and Template Haskell do offer some amount of static
typing \cite{conf/dagstuhl/CzarneckiOST03}, MetaOCaml\footnote{and
  MetaML.} offers stronger guarantees.  As well, since MetaOCaml is
implemented on top of the very solid Ocaml language, one accrues
additional benefits.

While it is theoretically feasible in a strings-based
generative system to re-parse the generative code to inspect it, such
an approach rapidly degenerates into complete chaos.  In languages
that natively offer reflection and reification (either at no cost, 
such as in LISP and Scheme, or at a modest cost in languages like 
Maple), this is actually feasible, even easy.  But this is fragile
and error-prone, since the languages offer little support for catching
errors early.

In MetaOCaml, generated code cannot be examined and optimized at all. It must
be generated just right.  For simple code, this is relatively easy
(see \cite{TahaThesis} for many examples).  For more complex examples,
new techniques are necessary.  \cite{KiselyovTaha} introduces one
such tool: abstract interpretation.  But more problems remain:
\jacques{can't find this 'Padua' paper at the 2004 Haskell Workshop 
you refer to}: 
\begin{itemize}
    \item Generating names in the presence of ``if'', ``while'', and other
        control operators. \cite{KiselyovTaha} showed how to solve
        this problem in simple cases.
    \item Making continuation-passing style (CPS) code clear.  Many
        authors understandably shy away from CPS code as it quickly
        becomes unreadable.  But this is needed for proper name
        generation.
    \item Compositionality of various interleaving aspects, as well
        as expressing dependencies among them.  For example, if we
        choose to compute the determinant of a matrix while performing
        Gaussian Elimination, this requires inserting code in 
        several specific places in the algorithm.  The method to
        compute this determinant crucially depends on whether 
        a division-based or a fraction-free Gaussian Elimination is
        used.
    \item As mentionned above, there are dependencies between
        certain algorithmic choices and properties of the
        underlying domain, and we would like to statically guarantee
        that inappropriate combinations cannot be chosen.  For instance,
        one should not attempt to use full division when dealing
        with matrices of exact integers, nor is it worthwhile to use
        full pivoting on a matrix of rational integers ($\mathbb Q $).
\end{itemize}
The question we want to answer is:
Can we achieve all of the abstractions needed for a flexible 
Gaussian Elimination, and can
we eliminate all of that abstraction's overhead in the generated code?
Of course, we want to do this without using
intensional code analysis -- that is, maintaining the
guarantees of MetaOCaml and using a purely generative approach.

Maintaining generativity is not just an interesting exercise:
one gets a stronger equational theory \cite{Taha2000}, and avoids
the danger of creating unsoundness \cite{TahaThesis}.  It should
also be remarked that in general, intensional code analysis
\cite{Pueschel:05,Kennedy01Telescoping,dongarra7,Veldhuizen:2004} 
essentially requires
one to insert both an optimizing compiler and an automated theorem
proving system into the code generating system.  While this is
potentially extremely powerful and an exciting area of research,
it is also extremely complex, which means that it is currently more
error-prone and difficult to ascertain the correctness of the 
resulting code.

This work makes the following contributions:
\begin{itemize}
    \item orderly aspect weaving
    \item mdo notation
    \item use of monads for meta-programming
    \item use of generative functors
    \item use of type constraints to ensure correctness
\end{itemize}

The rest of this paper is structured as follows:
The next section introduces code generation in MetaOCaml, the problem
of name generation, and how continuation-passing style is a general
tool for solving that problem.  Section~\ref{monadicnotation} introduces
the monad that greatly simplifies the task of dealing with
possibly effect-inducing CPS code combinators.  Section~\ref{functors}
shows how one can use the advanced module system of OCaml, namely its
parametrized modules, to encode all of the aspects of the Gaussian
Elimination algorithm family in completely separate, independent modules.
We then briefly discuss related work in section~\ref{related}, outline
some work still to do in section~\ref{future}, and finally draw some 
conclusions in section~\ref{conclusion}.  As an appendix, we show
the details of the syntax extension for monadic programming in Ocaml
and MetaOCaml.

The first author wishes to thank Cristiano Calgano for his help in
adapting camlp4 for use with MetaOCaml.

\section{CPS and the problem of generating names}\label{CPS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% And now, a better -- and shorter -- write-up.

The essence of our approach is assembling the code of the program by
combining instances of various aspects together. In the case of
Gaussian Elimination, the algorithmic aspects are pivot determination, 
swapping of
rows and columns, determinant tracking, etc. A primitive code
generator yields a code expression.  A combinator takes (complete and
well-typed) code expressions and combines them into a composite expression
(which is also statically guaranteed to be well-formed and well-typed).

We will be using MetaOcaml which, as an instance of a multi-stage
programming system \cite{TahaThesis}, provides exactly the needed
features: to construct a code expression, to combine them, and to
execute them.

For example, the following is probably the simplest code generator:
\begin{code}
let one = .<1>.
\end{code}

In the floating-point instance of that aspect, we would have
|.<1.0>.| instead. We use MetaOCaml brackets |.<...>.| to generate
code expressions -- in other words, to construct future-stage
computations. The simplest combination is inlining:
\begin{code}
let plus x y = .<.~x + .~y>.
\end{code}
Here we see the second component of MetaOCaml -- escapes |.~|. They
let us perform an immediate code generating computation while we are
building the future-stage computation (code expression). In the
example above, the immediate computation |.~x| is trivial -- obtaining
the value of the variable |x|, which should be a code expression, and
inlining it. The following shows an example of using the simple
generators:

\begin{code}
let simplest_code =
  let gen x y = plus x (plus y one) in
  .<fun x y -> .~(gen .<x>. .<y>.)>.
\end{code}

Here, the immediate computation |gen .<x>. .<y>.| is less trivial --
it includes the evaluation of the function |gen|, which in turn
applies the function |plus|. The function receives code expressions
|.<x>.| and |.<y>.| as arguments. At the generating stage, we can
manipulate code expressions as (opaque) values. The function |gen| returns a
code expression, which will be inlined in the place of the
escape. MetaOCaml can print out code expressions, so we can see the
final generated code:\\
  |.<fun x_1 -> fun y_2 -> (x_1 + (y_2 + 1))>.|

We must point out that |gen x y| looks like a combinator; its body 
involves several applications of the function |plus|. When
we look at the generated code however, we see no traces of the function
|plus| and no such administrative applications. Instead, we see the
fully inlined expression. These administative applications are done at
the code generation stage.

The final MetaOCaml feature, |.!| (pronounced ``run'') can be used to
execute the code expression: |.! simplest_code| is a function of two
integers, which we can apply: |(.! simplest_code) 1 2|. The original
|simplest_code| is not the function on integers -- it is a code
expression.

To see the benefit of code generation, we notice that we can easily
parameterize our code:

\begin{code}
let simplest_param_code plus one =
  let gen x y = plus x (plus y one) in
  .<fun x y -> .~(gen .<x>. .<y>.)>.
\end{code}

\noindent and use it to generate code that operates on integers, floating point
numbers or booleans -- in general, any domain that implements |plus|
and |one|:
\begin{code}
let plus x y = .<.~x +. .~y>. and one = .<1.0>. in
  simplest_param_code plus one
let plus x y = .<.~x || .~y>. and one = .<true>. in
  simplest_param_code plus one
\end{code}
Running the former expression yields the function on |float|s, whereas
the latter expression is the code expression for a boolean function.

This clearly shows the separation of concerns, namely of that for domain
operations.

Let us consider a more complex expression:
\begin{code}
let param_code1 plus one =
  let gen x y = plus (plus y one) (plus x (plus y one)) in
  .<fun x y -> .~(gen .<x>. .<y>.)>.
\end{code}

We notice two occurrences of |plus y one|. Depending on the
implementation of |plus| that may be quite a complex computation, and
so we would rather not do it twice. We may be tempted to rely on the
compiler's common-subexpression elimination optimization. We would
rather not rely on a ``sufficiently smart compiler'': when the
generated code is very complex, the compiler may overlook common
subexpressions.  Or the subexpressions may occur in an imperative
context where the compiler might not be able to determine if lifting
them is sound. So, being conservative, the optimizer will leave the
duplicates as they are. We may attempt to write

\begin{code}
let param_code1 plus one =
  let gen x y = 
     let ce = (plus y one) in  plus ce (plus x ce) in
  .<fun x y -> .~(gen .<x>. .<y>.)>.
\end{code}

However, the result of |param_code1 plus one|, which prints as\\
|.<fun x_1 -> fun y_2 -> ((y_2 + 1) + (x_1 + (y_2 + 1)))>.|,
still exhibits duplicate sub-expressions. Our |let|-insertion
optimization saved the computation at the generating stage. 
We need a combinator that inserts the |let| expression in the
generat\emph{ed} code. We need a combinator |letgen| to be used
as\\|let ce = letgen (plus y one) in plus ce (plus x ce)| 
yielding the code like |.<let t = y + 1 in t + (x + t)>.|
But that seems impossible because |letgen exp| has to generate
the expression\\|.<let t = exp in body>.| although |letgen| does not have
the |body| yet. The body needs a temporary identifier |.<t>.| that
is supposed to be the result of |letgen| itself. 
Certainly |letgen| cannot generate only part of a let-expression,
without the |body|,  as all generated expressions in
MetaOCaml are well-formed and complete.

The key is to use continuation-passing style (CPS). Its benefits were first
pointed out by \cite{Bondorf:92} in the context
of partial evaluation, and extensively used by \cite{KiselyovTaha} for code
generation.

\begin{code}
let letgen exp k = .<let t = .~exp in .~(k .<t>.)>.
let param_code2 plus one =
  let gen x y k = letgen (plus y one) (fun ce -> k (plus ce (plus x ce)))
  and k0 x = x
  in .<fun x y -> .~(gen .<x>. .<y>. k0)>.
\end{code}

Now, |param_code2 plus one| gives us the desired code\\
|.<fun x_1 -> fun y_2 -> let t_3 = (y_2 + 1) in (t_3 + (x_1 + t_3))>.|.

\section{Monadic notation, making CPS code clear}\label{monadicnotation}

Comparing the code that did let-insertion at the generating stage\\
|let ce = (plus y one) in  plus ce (plus x ce)|\\
with the corresponding code inserting let at the generated code stage\\
|letgen (plus y one) (fun ce -> k (plus ce (plus x ce)))|\\
clearly shows the difference between the direct-style and CPS code.
What was |let ce = init in ...| in direct style became
|init' (fun ce -> ...)| in CPS. For one thing, |let| became
inverted. For another, what used to be an expression that yields
a value, |init|, now became an expression that takes an extra argument,
the continuation, and invokes it. The differences look negligible in
the above example. In larger expressions with many let-forms, the
number of parentheses around |fun| increases, the need to add and
then invoke the |k| continuation argument become increasingly annoying. The
inconvenience is great enough for some people to explicitly avoid CPS
or claim that numerical programmers (our users) cannot or will not
program in CPS. Clearly a better notation is needed.

The |do|-notation of Haskell \cite{Haskell98Report} shows that it is possible
to write essentially CPS code in conventional-looking style. The
|do| notation is the notation for writing monadic code (cite Moggi?).
The benefit of monadic notation is not that it can represent CPS (cite
Filinski), but it helps in composability by offering to add different
layers of effects (state, exception, non-determinism, etc) to the
basic monad (cite Hudak, Peyton-Jones) in a controlled way.

A monad (cite Moggi?) is an abstract datatype representing
computations that yield a value and may have an \emph{effect}.
The datatype must have at least two operations, |return| to build
trivial effect-less computations and |bind| for combining
computations. These operations must satisfy \emph{monadic laws}:
|return| being the left and the right unit of |bind| and |bind| being
associative. Figure~\ref{ourmonad} defines the monad used throughout
the present paper and shows its implementation.

\begin{figure*}\label{ourmonad}
\begin{code}
type ('v,'s,'w) monad = 's -> ('s -> 'v -> 'w) -> 'w
let ret a = fun s k -> k s a
let bind a f = fun s k -> a s (fun s' b -> f b s' k)
let fetch s k = k s s
let store v s k = k (v::s) ()

let k0 s v = v
let runM m = m [] k0

let l1 f = fun x -> mdo { t <-- x; f t}
let l2 f = fun x y -> mdo { tx <-- x; ty <-- y; f tx ty}

let retN a = fun s k -> .<let t = .~a in .~(k s .<t>.)>.

let ifL test th el = ret .< if .~test then .~th else .~el >.
let ifM test th el = fun s k ->
  k s .< if .~(test s k0) then .~(th s k0) else .~(el s k0) >.
\end{code}
\caption{Our monad}
\end{figure*}

Our monad represents two kinds of computational effects: reading and
writing a computation-wide state, and control effects. The latter are
normally associated with exceptions, forking of computations, etc. --
in general, whenever a computation ends with something other than
invoking its natural continuation in the tail position. In our case
however the control effects manifest themselves as code generation.

In Figure~\ref{ourmonad}, the monad is implemented as a function of two
arguments: the state (of type |s|) and the continuation. The
continuation receives the current state, the value (of the type |v|) and
yields the answer of the type |w|.  The monad is polymorphic over all
these three type parameters.  Other implementations are
possible. Except for the code in Figure~\ref{ourmonad}, the rest of our code
treats the monad as a truly abstract data type. The implementation of basic
monadic operations |ret| and |bind| are conventional. It is easy to
see that the monadic laws are satisfied.  Other monadic operations
construct computations that do have effects.  Operations |fetch| and
|store v| construct computations that read and write the state. In our
case the state is a list (of polymorphic variants), which models an
open discriminated union, as we shall see later.

The operation |retN a| is the let-insertion operation, whose simpler
version we called |letgen| earlier. It is the first computation with
a control effect: indeed, the result of |retN a| is \emph{not} the
result of invoking its continuation |k|. Rather, its result is a |let|
code expression. \oleg{Mention similarity with reset?}
\jacques{yes}

Finally, |runM| runs our monad, that is, performs the computation of
the monad and gets the result, which in our case is the code
expression. We run the monad by passing it the initial state and the
initial continuation |k0|. We can now re-write our |param_code2|
example of the previous section as

\begin{code}
let param_code3 plus one =
  let gen x y = bind (retN (plus y one)) (fun ce -> 
                ret (plus ce (plus x ce)))
  in .<fun x y -> .~(runM (gen .<x>. .<y>.))>.
\end{code}
% param_code3 plus one;;

That does not seem like much of an improvement. But here we can introduce the
do notation (Appendix A), which is patterned after the do-notation of Haskell.

\begin{code}
let param_code4 plus one =
  let gen x y = mdo { ce <-- retN (plus y one);
                      ret (plus ce (plus x ce)) }
  in .<fun x y -> .~(runM (gen .<x>. .<y>.))>.
\end{code}

The function |param_code4| is totally equivalent to |param_code3| --
the |mdo|-notation is just syntactic sugar. In fact, the camlp4
preprocessor (with our extension) will transform |param_code4|
into exactly |param_code3|. And yet, |param_code4| looks far more
conventional, as if it were indeed in direct style.

We can write operations that generate code other than let-statements,
e.g., conditionals: see |ifL| in Figure~\ref{ourmonad}. The function |ifL|, 
albeit straightforward, is not as general as we wish: its argument are
already generated pieces of code rather than monadic values. So, we
may want to ``lift it'':
\begin{code}
let ifM' test th el = mdo {
  testc <-- test;
  thc   <-- th;
  elc   <-- el;
  ifL testc thc elc}
\end{code}

\noindent We can define functions |l1|, |l2|, |l3| (analogues of |liftM|,
|liftM2|, |liftM3| of Haskell) to make such a lifting generic.
We soon notice the need for another |ifM| function, with the same
interface (see ifM in Figure~\ref{ourmonad}). The difference between them is
apparent from the following:
\begin{code}
let gen a i = ifM' (ret .<(.~i) >= 0>.) (retN .<Some (.~a).(.~i)>.)
                           (ret .<None>.)
 in .<fun a i -> .~(runM (gen .<a>. .<i>.))>.
;;
.<fun a_1 -> fun i_2 ->
    let t_3 = (Some (a_1.(i_2))) in if (i_2 >= 0) then t_3 else (None)>.
\end{code}
\noindent whereas if we replace |ifM'| with |ifM|, the result will be
\begin{code}
.<fun a_1 -> fun i_2 ->
    if (i_2 >= 0) then let t_3 = (Some (a_1.(i_2))) in t_3 else (None)>.
\end{code}

The difference is apparent: in the code with |ifM'|, the let-insertion
happend \emph{before} the if-expression, that is, before the test that
the index |i| is positive. That is, if |i| turned out
negative, |a.(i)| will generate an out-of-bound array access
error. On the other hand, the code with |ifM| accesses the array only
when we have verified that the index is non-negative. This example
makes it clear that the code generation (such as the one in |retN|) is 
truly an effect and we have to be clear about the sequencing of
effects, when generating control constructions such as conditionals.
The form |ifM| handles the such effects correctly. In our code, we
need similar operators for other Ocaml control forms: for generating
case-matching statements and for- and while-loops. To make writing
such code general and independent of the implementation of the monad,
we define a function XXX. As the name implies, we do use control
effect implemented via essentially delimited continuations (cite?).
\jacques{yes!}

\section{Functors}\label{functors}

Now that we know that we have the appropriate basic techniques for
writing our code generators, we need to worry about modularizing them
appropriately.  Here we can really take advantage of the fact that
MetaOCaml is a staged language: it is not really relevant how much
computation happens at staging time, as long as the generated code
is correct and efficient.  In other words, we can use any abstraction
mechanisms we want to structure our code generators, as long as none
of those abstractions infiltrate the generated code.

While the Object-Oriented Design community has acquired an extensive
vocabulary for describing modularity ideas, the guiding principles for
modular designs has not changed since they were first articulated by
Parnas~\cite{journals/cacm/parnas72a} and Dijkstra~
\cite{EWD:EWD447}: information hiding and separation of concerns.  To
apply these principles to the study of Gaussian Elimination, we need
to understand what changes between different implementations, and 
what concerns need to be addressed.  We also need to study the degree
to which these concerns are independent.

A careful study of Gaussian Elimination \cite{carette04} shows that
(at least) the following variations occur:
\begin{enumerate}
	\item \textbf{Domain}: In which (algebraic) domain do the matrix
		elements belong to.  Sometimes the domains are very specific
		($\mathbb{Z}, \mathbb{Q}, \mathbb{Z}_p, 
		\mathbb{Z}_p\left[\alpha_1,\ldots,\alpha_n\right], 
		\mathbb{Z}\left[x\right]$, $\mathbb{Q}\left(x\right)$, 
		$\mathbb{Q}\left[\alpha\right]$, and floating point numbers for 
		example), while in other cases the domains were left generic,
		namely for elements of a field,
		multivariate polynomials over a field, elements of a formal Ring
		with possibly undecidable zero-equivalence, or elements of a 
		Division Ring.  In the roughly 85 pieces of code surveyed,
		20 different domains were encountered.
	\item \textbf{Container}: Whether the matrix
		is represented as an array of arrays, a one-dimensional array,
		a hash table, a sparse matrix, etc., and
		whether indexing is done in C or Fortran style.  Additionally,
		if a particular representation had a special mechanism for efficient
		row exchanges, this is sometimes used.  
	\item \textbf{Output choices}:  Whether just the reduced matrix, or
		additionally the rank, the determinant, and the
		pivoting matrix are to be returned.
		It is worthwhile noting that in the larger algorithm family,
		routines like Maple's \texttt{LinearAlgebra:-LUDecomposition} have
		up to $2^6 + 2^5 + 2^2 = 100$ possible outputs.
	\item \textbf{Fraction-free}: Whether the Gaussian Elimination
		algorithm is allowed to use unrestricted division, or only
		exact (remainder-free) division.
	\item \textbf{Pivoting}: Whether to use no pivoting at all, only
		column-wise pivoting, or full pivoting.
	\item \textbf{Augmented Matrices}: Whether we are doing GE on
		a full matrix, or only a restricted number of columns, while
		doing elimination on the full matrix.  Note that we currently
		do not treat this aspect in our code.
\end{enumerate}
\noindent In addition to the above variations, there are two aspects that 
recur frequently:
\begin{enumerate}
	\item \textbf{Length measure}:  For stability reasons
		(whether numerical or coefficient growth), if a domain possesses
		an appropriate length measure, this is sometimes used to choose
		an ``optimal'' pivot for row exchanges.
	\item \textbf{Normalization and zero-equivalence}: Whether the 
		arithmetic operations of the domain at hand gives results in 
		normalized form, and whether a specialized zero-equivalence 
		routine needs to be used.
\end{enumerate}
\noindent These are separated out from the others are they are cross-cutting
concerns: in the case of the length measure, then a property of the domain
will influence the method used to do pivoting \textbf{if} pivoting is to be
performed.

The simplest parametrization is to make the domain abstract.  As it 
turns out, we need the following to exist in our domains:  $0$, $1$,
$+$, $*$, (unary and binary) $-$, some kind of division, normalization,
and potentially a relative size measure.  We definitely need at least
\emph{exact} division to exist, but if full division is available, than
can be used as well.  Normalization is frequently trivial for domains
where elements have unique representations, but in many applications
we have domains (like polynomials in sum-of-product form) where this
is not the case.  This leads us to defining a |DOMAIN| type as
\begin{small}
\begin{code}
module type DOMAIN = sig
  type v
  type kind (* Field or Ring ? *)
  type 'a vc = ('a,v) code
  val zero : 'a vc
  val one : 'a vc
  val plus : 'a vc -> 'a vc -> ('a vc, 's, 'w) monad
  val times : 'a vc -> 'a vc -> ('a vc, 's, 'w) monad
  val minus : 'a vc -> 'a vc -> ('a vc, 's, 'w) monad
  val uminus : 'a vc -> ('a vc, 's, 'w) monad
  val div : 'a vc -> 'a vc -> ('a vc, 's, 'w) monad
  val better_than : ('a vc -> 'a vc -> 
      (('a,bool) code, 's, 'w) monad) option
  val normalizerf : (('a,v -> v) code ) option
  val normalizerg : 'a vc -> 'a vc
end 
\end{code}
\end{small}
\noindent  The first item to notice is that the types above are
generally lifted twice: once from the value domain |v| to the code
domain |'a vc|, and once more from values to monadic computations
|('a vc, 's, 'w) monad|.  Since there are times where we don't want
to redundantly walk a whole matrix to apply normalization when 
normalization is not necessary, |normalizerf| is optional.

For example, when lifted to a |DOMAIN|, the type |int| becomes
\begin{small}
\begin{code}
module IntegerDomain = struct
    type v = int
    type kind = domain_is_ring
    type 'a vc = ('a,v) code
    let zero = .< 0 >.  
    let one = .< 1 >. 
    let plus x y = ret .<.~x + .~y>. 
    let times x y = ret .<.~x * .~y>.
    let minus x y = ret .<.~x - .~y>.
    let uminus x = ret .< - .~x>.
    let div x y = ret .<.~x / .~y>. 
    let better_than = Some (fun x y -> ret .<abs .~x > abs .~y >. )
    let normalizerf = None 
    let normalizerg = fun x -> x
end
\end{code}
\end{small}
\noindent While the |DOMAIN| type may have looked daunting to some, 
the implementation is quite straightforward.  It is equally simple to 
make implementations for |float| and for |Num.num| (arbitrary precision
exact rational numbers).  With scarcely more work (as these are not 
built-in and the arithmetic must be implemented), $\mathbb{Z}_p$,
$\mathbb{Z}\left[x\right]$, $\mathbb{Q}\left(x\right)$ and more
complex domains can be done as well.

Parametrizing by the kind of container used to represent a matrix is
almost as straightforward.  We choose to make the container parametric
over a |DOMAIN| by making containers functors from a |DOMAIN| module
to the actual implementation of a container.  Our containers are
explicitly 2-dimensional, and we provide functions |dim1| and |dim2|
to extract each.  We also need |get| and |set| functions -- these come
in both a monadic flavour as well as a direct code-generation version.
For convenience and efficiency, several additional functions are also
required: a way to |copy| a container, a method (|mapper|) to map an
optional function over the elements of a container, as well as methods
to swap rows and columns.  The inclusion of these functions in the 
signature of all containers makes it simpler to optimize the relevant
functions depending on the actual representation of the container
while not burdening the users of containers with efficiency details.
As an example, below is an implementation of a matrix as an
array of rows (themselves arrays).  This implementation clearly shows
how swapping rows is more efficient than swapping columns in this
representation.  We have not (yet) implemented the further potential
optimization wherein if it is known that part of two rows/columns are already
known to be equal (in fact to |Dom.zero| in the case of GE) and
row/column exchange needs to be performed element-wise, then certain
swaps are unecessary.

\begin{small}
\begin{code}
module GenericArrayContainer(Dom:DOMAIN) =
  struct
  type contr = Dom.v array array (* Array of rows *)
  type 'a vc = ('a,contr) code
  type 'a vo = ('a,Dom.v) code
  let get' x n m = .< (.~x).(.~n).(.~m) >.
  let get x n m = ret (get' x n m)
  let set' x n m y = .< (.~x).(.~n).(.~m) <- .~y >.
  let set x n m y = ret (set' x n m y)
  let dim2 x = .< Array.length .~x >.       (* number of rows *)
  let dim1 x = .< Array.length (.~x).(0) >. (* number of cols *)
  let mapper g a = match g with
      | Some f -> .< Array.map (fun x -> Array.map .~f x) .~a >.
      | None   -> a
  let copy = (fun a -> .<Array.map (fun x -> Array.copy x) 
                       (Array.copy .~a) >. )
  (* this can be optimized with a swap_rows_from if it is known that
     everything before that is already = Dom.zero *)
  let swap_rows_stmt a r1 r2 =
      .< let t = (.~a).(.~r1) in
         begin 
             (.~a).(.~r1) <- (.~a).(.~r2);
             (.~a).(.~r2) <- t
         end >.
  let swap_cols_stmt a c1 c2 = .< 
      for r = 0 to .~(dim2 a)-1 do
          let t = (.~a).(r).(.~c1) in
          begin 
              (.~a).(r).(.~c1) <- (.~a).(r).(.~c2);
              (.~a).(r).(.~c2) <- t
          end
      done  >.
end
\end{code}
\end{small}
\noindent  Implementing other containers based on flattened vectors,
Fortran-style access, |Bigarray| or |Bigarray.Array2| is also possible.
Naturally, for proper use of |Bigarray| with specific types, it is 
important that the |DOMAIN| type matches -- but luckily the compiler 
will prevent us from making such a mistake.

The use of a |functor| for making a container parametric is fairly 
straightforward.  It gets more interesting when one turns to the aspect
of what to return from the GE algorithm.  One could create an algebraic
data type (as was done in \cite{carette04}) to encode the various
choices; but this is wholly unsatisfying as we know that for any single
use, only one of the choices is ever possible, yet any routine which
calls the generated code must deal with these unreachable options.
Instead we use a module type with an \emph{abstract} type |res| for the
result type; the actual result type is then made by instantiating
this module type with different modules that decide what will be 
output: the matrix, the matrix and the rank, the matrix and the
determinant, the matrix, rank and determinant, and so on.  More
precisely, below we show this type and one instantiation.

\begin{small}
\begin{code}
module type OUTPUT = sig
  type contr
  type res
  module D : DETERMINANT
  module R : RANK
  module P : TRACKPIVOT
  val make_result : ('a,contr) code -> 
    (('a,res) code,
     [> `TDet of 'a D.lstate | `TRan of 'a R.lstate | `TPivot of 'a P.lstate]
       list,
     ('a,'w) code) monad
end

module OutDetRank(Dom:DOMAIN)(C: CONTAINER2D)
    (Det : DETERMINANT with type indet = Dom.v and type outdet = Dom.v)
    (Rank : RANK) =
  struct
  module Ctr = C(Dom)
  type contr = Ctr.contr
  type res = contr * Det.outdet * int
  module D = Det
  module R = Rank
  module P = DiscardPivot
  let make_result b = mdo {
    det  <-- D.fin ();
    rank <-- R.fin ();
    ret .< ( .~b, .~det, .~rank ) >. }
end
\end{code}
\end{small}
\noindent It is worthwhile noting that while the |OUTPUT| type is
relatively simple, instantiating it requires a rather complex functor
with some additional type constraints.  While we will explain some of
the details of the above code later, right now it suffices to notice
that |res| above is a 3-tuple with type |contr * Det.outdet * int|,
where in another instance this type is simply |contr|, and these are
formally incomparable types.

As is apparent from the output choices, several different quantities
\emph{may} need to be tracked in a particular GE implementation.
We therefore need to be able to conditionally generate code-variables,
and weave in corresponding tracking code for each of these quantities.
To be precise, we may need to keep track (independently) of the rank,
the determinant and the permutation list.  These pieces of data then
become part of the \emph{state} that is tracked by our monad.  To have
all this choice when needed, and yet have our code be modular and
composable as well as ensuring that the generated code does not 
contain any abstraction artifacts, it is important to make this 
state modular.  As this discussion is rather abstract, an example
will likely illustrate our point more clearly: we will show the 
code for tracking the determinant.

\begin{small}
\begin{code}
module type DETERMINANT = sig
  type indet
  type outdet
  type tdet = outdet ref
  type 'a lstate
  val decl : unit -> 
    (unit, [> `TDet of 'a lstate ] list, ('a,'b) code) monad
  val upd_sign : unit -> 
    (('a,unit) code, [> `TDet of 'a lstate ] list, ('a,'b) code) monad
  val zero_sign : unit -> 
    (('a,unit) code, [> `TDet of 'a lstate ] list, ('a,'b) code) monad
  val acc : ('a,indet) code -> 
    (('a,unit) code, [> `TDet of 'a lstate ] list, ('a,'b) code) monad
  val get : unit ->
    (('a,tdet) code, [> `TDet of 'a lstate ] list, ('a,'b) code) monad
  val set : ('a,indet) code -> 
    (('a,unit) code, [> `TDet of 'a lstate ] list, ('a,'b) code) monad
  val fin : unit -> 
    (('a,outdet) code, [> `TDet of 'a lstate ] list, ('a,'b) code) monad
end
\end{code}
\end{small}
\noindent  Determinant tracking is quite complex: we need to 
declare that we are tracking the determinant (|decl|), have some functions
for tracking the sign of the determinant (|upd_sign|,|zero_sign|),
|set|, |get|, accumulate (|acc|) unsigned values for the determinant, as 
well as returning a final value (|fin|) for the determinant.  In some cases,
like for float matrices where we do not return the determinant, all of
the above should generate no code at all, while for integer matrices where
we want the determinant as part of the output, each of these need to be
implemented.

\jacques{I am stopping my description of state here -- what I have
written above is quite dull, and not making the point very well.
I will need to revisit this}

Making state modular and composable: several aspects can contribute
to the state, without interference (show: Rank, Det, Permutation
aspects).
What is state and why we need it: making aspects interleave: use det
as an example. State as an open record. List of polymorphic variants
appeared to be the easiest way to implement such a union, in a pure
functional way.

Expressing domain constraints in functors (Div and integer domain).
Show the type error that occurs in improper instantiation.

Show several examples (e.g., code with NullPivot and nodet: no traces
of either det nor pivoting). Then the code with RowPivoting, Det and
float domain. \jacques{which would really be great if we had a lot 
more space, but I think we're already long, so something is going to
have to be cut.  Maybe not this, but something!}

\section{Related work}\label{related}

Note that the monad is similar to the one used in \cite{KiselyovTaha}.
However, the latter work used only |retN| of all monadic operations,
and used fixpoints (for performing iterations at the generation time).
In this paper we do not use monadic fixpoints (because the generator
is not recursive) but we make extensive use of monadic operations for
generating conditional and looping operations.

\jacques{What do we consider \emph{related work} here?  Do we essentially gloss 
over related work in numerical software, code generation in general,
aspect-oriented stuff?  We don't want this section to be too long, but
not so short that we make it seem like we're inventing more than we
really are}

\section{Future work}\label{future}
connections with delimited continuations: making notation
more direct-style and potentially clearer.  More syntactic sugar
would help a lot here.

More aspects that can be handled: Input variations (augmented
matrices). Fit into the larger program family, where there are
more aspects still.

Making all of the algorithmic code be completely independent of 
MetaOCaml markup and isolate all the code generation to a select
few 'base' modules and functors.  This should allow for simpler
debugging, where one can choose to instantiate either a direct-style
set of base modules for debugging, or a code-generation style set
of base modules for generating optimized code.

\section{Conclusion}\label{conclusion}

Mention the mdo notation. Mention it into the introduction.

Relation with aspect-oriented code: in AspectJ, aspects are (comparatively)
lightly typed, and are post-facto extensions of an existing piece of
code.  Here aspects are weaved together ``from scracth'' to make up a
piece of code/functionality.  One can understand previous work to be
more akin to dynamically typed aspect weaving, while we have started
investigating statically typed aspect weaving.

\bibliography{metamonads}
\bibliographystyle{plain}

\section{Appendix A}
The do notation for Ocaml.
We use camlp4. Show the grammar. Explain the difficulty: although
nice, the do notation cannot actually be expressed in a LR(n) grammar.
Perhaps give the command line how to load the camlp4 module into
ocaml. Point to the code.
\end{document}

\omitnow{
In this paper, we are concerned with building and composing
generators. We start with the code expressing the algorithm (in
OCaml), and, adding MetaOCaml annotations, turn that code into a
\emph{code expression} that will generate the code when runs. We then
modify that code expression to add various parameters and aspects, to
yield the complete generator.

Here we illustrate the MetaOCaml constrcutions and code generators.

Turning code into the code expression is relatively straightforward
until we come to binding constructs such as |let|. For example, here
is a piece of code that defines a function that searches a given array
|arr| for the maximal (in absolute value) element, and the index of
that element and the found absolute value. This code is representative
of pivot determination. The code uses mutations and destructive
updates -- partly for illustrative purposes and mainly because many
linear algebra algorithms are expressed via an imperative code (so it
is less error-prone to translate the ``canonical'' algorithms with
minimal changes).

\begin{code}
let fp = fun arr n ->
   let pv = ref (0,arr.(0)) in
   for i = 1 to n-1 do
     if abs_float arr.(i) > snd !pv then
        pv := (i, abs_float arr.(i))
   done;
   !pv
\end{code}

This function has the type |float array -> int -> int * float|. By
enclosing the code into MetaOCaml brackets, we obtain the simplest
generator expression.

\oleg{later on, we can write the following just let fpc = .<...>.}

\begin{code}
let fpc = .<fun arr n ->
   let pv = ref (0,arr.(0)) in
   for i = 1 to n-1 do
     if abs_float arr.(i) > snd !pv then
        pv := (i, abs_float arr.(i))
   done;
   !pv>.
\end{code}

This value has the type |('a, float array -> int -> int * float) code|.
Here |'a| type parameter is needed (ask Walid for the reference).
The main difference between |fp| and |fpc| is that |fp| is a function,
which we can immediately use as \verb^fp [|1.0; -4.; 3.|] 3^. In
contrast, |fpc| is not a function: it is a \emph{code} value that denotes
that function. We would say that the code expression is the
``second-stage'' value. The ability to manipulate code expressions as
regular values is the distinguished fature of MetaOCaml.
We can ``compile'' the code by using MetaOCaml |run| command, or |.!|:
|.! fpc|. The result is equivalent is |fp|. Although |fpc| doesn't
seem to do much, it is already useful as we can use offshoring (cite)
to translate |fpc| into C or Fortran code (which we can dynamically
link into the running MetaOCaml code -- or use separately).

The code |fpc| obviously lacks generality: the name |abs_float|
implies that that code deals with |float| arrays. We wish to handle
arrays in various domains, integer, rationals, etc. We can make the
code generic by parametrizing over the |abs| function:

\begin{code}
let fpca absf = .<fun arr n ->
   let pv = ref (0,arr.(0)) in
   for i = 1 to n-1 do
     if absf arr.(i) > snd !pv then
        pv := (i, absf arr.(i))
   done;
   !pv>.
\end{code}

and so |fpca abs_float| will manipulate |float array| and |fpca abs|
will do for the integer arrays. Here we used the value from the first
stage, |absf|, to incorporate into the code at the second stage:
croos-stage persistent values.

We should now consider the function |fpca| in a browder context: of a
function that computes the pivot and then swaps the pivot withe 0-th
element of the array:

\begin{code}
let upv = .<fun arr n ->
   let pv = .~(fpca abs_float) arr n in
   let i = fst pv in
   (if not (i = 0) then
      let t = arr.(0) in
      begin arr.(0) <- arr.(i); arr.(i) <- t end);
   arr>.
\end{code}

Here, we see the third MetaOCaml operator, escape |.~|. It does ...
MetaOcaml can print out the code, which looks like
|.<fun arr n -> let pv = (fun arr n -> fpcs code) arr n in ... >.|
That is, the compiler inlined the code for fpca -- which is a
functional expression. The |upv| code shows the functional
application, |(fun arr n -> ...)| applied to arguments |arr n|. That
application will be executed at run-time, unless a sufficiently smart
compiler can inline it. As we stated, our aim is to not to rely on
post-generation optimizations and sufficiently smart compiler. We
strive to generate the efficient code ourselves. We can do that by
modifying the definition of |fpca| as

\begin{code}
let fpca absf arr n = .<
   let pv = ref (0,(.~arr).(0)) in
   for i = 1 to .~n-1 do
     if absf (.~arr).(i) > snd !pv then
        pv := (i, absf (.~arr).(i))
   done;
   !pv>.
\end{code}
\begin{code}
let upv = .<fun arr n ->
   let pv = .~(fpca abs_float .<arr>. .<n>.) in
   let i = fst pv in
   (if not (i = 0) then
      let t = arr.(0) in
      begin arr.(0) <- arr.(i); arr.(i) <- t end);
   arr>.
\end{code}
% (.! upv) [|1.0; -4.; 2.|] 3;;
We see that we pass code fragment |.<arr>.| as arguments to the
function. Now when we look at the code for |upv|, we see
|.<fun arr n -> let pv1 = let pv = ref (0,arr.(0)) in ...>.|
So the fpca code got really inlined.

The generator |fpca| shows one particular way of searching for pivot.
In the domain of exact rational numbers, for example, we would be
satisfied with the first non-zero element as a pivot, and so we would
write a different pivot-searching generator. The body of |upv| moves
the found pivot to the ``canonical'' location. Again, depending on the
container (a dense matrix, a sparse matrix, etc), there are different
ways of doing that. So, we would like to be able to ``compose'' pivot
generating function with the swap generating function. We would like
to write something like this

\begin{code}
let fpca absf arr n pv = .<
   for i = 1 to .~n-1 do
     if absf (.~arr).(i) > snd ! (.~pv) then
        (.~pv) := (i, absf (.~arr).(i))
   done>.
let swapper arr n pv = .<
   let i = fst !(.~pv) in
   if not (i = 0) then
      let t = (.~arr).(0) in
      begin (.~arr).(0) <- (.~arr).(i); (.~arr).(i) <- t end>.
let seqv a b c = .<begin .~a; .~b; .~c end>.
let pv_swap absf arr n =
   let pv = .<ref (0,(.~arr).(0))>. in
   seqv (fpca absf arr n pv)
        (swapper arr n pv)
        arr
let top =  .<fun arr n -> .~(pv_swap abs_float .<arr>. .<n>.)>.
\end{code}

The function |pv_swap| shows our ideal -- composing code gerenators
|fpca| and |swapper|, using a higher-order combinator `seqv'
(sequence of code fragments). We lifted the pivot declaration out of
old version of |fpca| so that both |fpca| and |swapper| could be
parameterized by the pivot (and |fpca| and similar function will deal
only with pivot determination). |top| puts it all together. The above
code represents the pattern we will be following for the rest of the
paper.

And it \emph{almost} works. It is type correct.  When we look at the
generated code for |top|, we notice the |(ref (0, arr_1.(0)))| is
inlined wherever pivot was expected. For example, the swapping part of
the code begins with |let i_3 = (fst (! (ref (0, arr_1.(0))))) in ...|. 
That is certainly not what we want. We meant for pivot searching
and swapping parts of the code to communicate via a shared mutable
state, variable |pv|. In the generated code, each piece of code gets
its own copy of |pv|. Rather then inline |(ref (0, arr_1.(0)))|
everywhere, we should have generated |let pv = ref (0,(.~arr).(0)) in ...|
and inlined |pv|. Incidentally, even in a pure functional code we
often wish to generate such a |let| statement so we can compute a
complex expression only once and use it value in several other
places. Again, we would rather not leave the job of detecting such
common subexpressions to the compiler (because in the complex code the
compiler may fail to detect the common subexpressions). The problem
with generating such |let| expressions is that we wish our geenrator
to return just |.<let pv = ref (0,(.~arr).(0)) in>.|, without the `body'
of the |let|, which we would fill in later. In languages which
generate code in the string form this is easy. We generate the string
|''let pv = ref (0,(.~arr).(0)) in''| and then append to it the string
representting the body, wcich would give us the complete |let|
statement. In MetaOCaml, we can't generate incomplete expressions, for
a good reason.

To solve the problem of writing a generator for a |let| expression 
when the generated body is not yet known, we use CPS.
}