### 2006-09-14 TOMS/paper/paper.tex
style files and rough outline of the TOMS paper

\documentclass[acmtoms]{acmtrans2m}

%\acmVolume{V}
%\acmNumber{N}
%\acmYear{YY}
%\acmMonth{M|

\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newdef{definition}[theorem]{Definition}
\newdef{remark}[theorem]{Remark}

\markboth{J. Carette and O. Kiselyov}{Our grand title}

\title{So what will our title be?}

\author{JACQUES CARETTE\\ McMaster University
    \and OLEG KISELYOV\\ FNMOC}

\begin{abstract} 
We describe statically typed generative techniques that enable one to write
generic yet highly efficient software.  We demonstrate the effectiveness
of this for various common scientific and symbolic programming tasks.

Furthermore, we also show how these techniques allow for significantly
more modularization than previous work, at no loss of program efficiency.
We are thus able to extract even more design aspects out of linear algebra
codes than previous work.  We are thus able to seamlessly bridge the gap
between numerical and symbolic linear algebra.  The end result is that
the core abstract algorithm becomes the program text for generating all
its specialization, and the various design decisions required to generate
these are made completely explicit.
\end{abstract}

\category{G.4}{Mathematical Software}{???}

\terms{Design,Performance,Languages}

\keywords{Abstraction, genericity, MetaOCaml, metaprogramming}

\begin{document}

\begin{bottomstuff} 
To be filled in later
\end{bottomstuff}

\maketitle

\section{Introduction}

Talk about how, at no loss of efficiency, one can pull out very abstract
algorithms out of countless variants.  This even allows creating new
variants at essentially no cost to the author of the code.  It can pull
out the real core of an algorithm and make visible all the choices that
can be made.  Generalizations then become significantly easier.

In fact, in some cases efficiency can even be gained. 
The easiest way is by taking out dynamic choices and making them static.
Another is to allow more structure (like sparsity) to be taken advantage
of in a higly modular manner; thus authors are more likely to have the
time to code special versions.  Also, various optimization techniques
(like loop unrolling) can be applied at no maintenance cost.

The order of the sections in the paper is not fixed yet.

\section{Related Work}
Must mention Boost (and lots of other C++ work).  Revisit old emails to
find more references.  Not clear if this goes here or towards the end.

\section{Problem}
Describe the problem solved.  Basically a highly typed way to extract
all the design information from code variants into modular code generators
that compose correctly \emph{by construction}.  Requires encoding as
much information as possible in types.  But also, since MetaOCaml is
stage, some information can be encoded in values available at
``compile'' time, and this information can be checked (again at compile
time) for correctness and consistency before any code is generated.
This leads one down partway to dependent types, which will not be 
explored in this particular work.

There are multiple goals:
\begin{itemize}
\item Generate correct code
\item Generate efficient code
\item Modular generator
\item Design-points discovery
\end{itemize}

\section{MetaOCaml}
For TOMS people, an introduction to MetaOCaml will really be needed.  A
comparison to meta-programming in other languages, especially C++, will
also be necessary.

\section{Methodology}
Describe the tools and techniques.  Repeat some of the ideas from the 
GPCE paper, add the new ones.

\section{Case studies}
Go through all our examples, in detail.  Different algorithms have 
different numbers of easily identifiable design points that can be
extracted.  But the same mechanism works accross them all.

\subsection{Gaussian Elimination family}
Describe the members of the family, all the various bits that can be
extracted out.  Some pieces are incompatible --- show how one insures
they cannot successfully be composed with each other (by construction).

The family is much larger than just GE, as it also includes LU and
in general GE-based linear solving.

\subsection{Linear ODE solving family}
Need to put more variants in the code to be able to talk about the 
family some more.

\subsection{Cholesky family}
Obvious question: does this belong here or in GE?  We'll need to 
actually code this to find out I think!

\section{Conclusion}
Always the last thing written?

\bibliographystyle{acmtrans}
\bibliography{paper} % rename this!

\begin{received}
Received Month Year; revised Month Year; accepted Month Year
\end{received}
\end{document}