### 2007-05-15 SCP2006/paper/ge.tex
rename SCP to be more precise

\documentclass[11pt]{elsart}

\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{times}
\usepackage{stmaryrd}

\usepackage[backref,colorlinks,bookmarks=true]{hyperref}

\begin{document}
\begin{frontmatter}
\title{Gaussian Elimination: a case study in efficient genericity with
MetaOCaml}
% --- Author Metadata here ---
\author{Jacques Carette\thanksref{nserc}}
\thanks[nserc]{Work supported in part by NSERC Discovery Grant XXX}
\ead{carette@mcmaster.ca}
\ead[url]{http://www.cas.mcmaster.ca/\textasciitilde carette}

% --- End of Author Metadata ---

\begin{abstract}
The Gaussian Elimination algorithm is in fact an algorithm family -
common implementations contain at least 6 (mostly independent)
``design choices''.  A generic implementation can easily be parametrized
by all these design choices, but this usually leads to slow and bloated
code.  Using MetaOCaml's staging facilities, we show how we can
produce a natural and type-safe implementation of Gaussian Elimination 
which exposes its design choices at code-generation time, so that these choices
can effectively be specialized away, and where the resulting code is 
quite efficient.
\end{abstract}

\begin{keyword}
MetaOcaml, linear algebra, genericity, generative, staging
\PACS code
\end{keyword}
\end{frontmatter}

\section{Introduction}

Gaussian Elimination is a fundamental algorithm in computational 
mathematics.  As it finds uses in many areas of mathematics, quite a
number of variants on the basic algorithm have been invented over
the years, but the basic algorithm persists through all variants.  
An analysis of the source code \footnote{Note to reviewers: for 
completeness, this paper includes source code samples in Maple
and the complete MetaOCaml implementation in appendices.  The
code which is deemed inessential for a final print version will
instead be made available on the Web}
for the library of Maple\footnote{Maple
is a registered trademark of Waterloo Maple Inc.}, a large
commercial computer algebra system, found 35 different implementations
of Gaussian Elimination (GE), and at least as many more implementations of
highly related algorithms (LU factorization, Cholesky factorization,
reduction to Smith Normal Form, linear equation solving, etc).  There
was little actual code duplication between these implementations, as they all
embodied a certain set of design decisions that were hard-coded.  As
these design decisions have a measurable effect on the actual shape of the
code, there was little obvious opportunity for abstraction.  However,
since the authors of the code knew that the underlying algorithm was
all the same, it would have been possible to use higher-order functions
to parametrize these choices.  On the other hand,
a number of these variants are rather esoteric, or depend on quite subtle
properties of the underlying domain at hand - certainly an average user
who knew they needed to use Gaussian Elimination could not be expected to
understand all these choices, and furnish the appropriate functions.
Thus it still necessary to expose this algorithm via different 
specialized interfaces.  But even when
there were parametrization opportunities, the ensuing loss of efficiency 
between the generic code versus the specialized code (up to a factor of 30,
in part due to the fact that Maple is a dynamically typed 
interpreted language, but even in a statically typed language, 
the overhead of an (indirect) function call, a likely pipeline stall
and cache miss(es) would produce similar overhead) 
naturally
reinforced the idea that all these versions need to be coded seperately.
This clearly leads to a maintenance nightmare.
Thus we seek to answer the question:
\begin{center}
\emph{Is it possible to write a generic yet efficient Gaussian Elimination
algorithm?}
\end{center}

\noindent In Maple itself, this question can be answered positively:
Maple has very good reflection and reification capabilities
(see~\cite{JoGoSe93} for definitions), with
full introspection of all programs being quite easy; in fact, these
capabilities compare favorably to those of LISP/Scheme~\cite{Abelson:1998:RRA}.
Maple's reflection/reification are somewhat more type-safe than 
Scheme's; like in Scheme it is not possible to reify procedures 
which are syntactic nonsense, but also some mild type invariants are
also guaranteed to hold, although the result may still be obvious
semantic nonsense.  We are more interested in the following more precise
question:

\begin{center}
\emph{Is it possible to write a generic, type-safe yet efficient 
Gaussian Elimination algorithm?}
\end{center}

To ensure type safety, it is necessary to move to a typed language; to
allow for efficiency, it is necessary to remove all the overhead
introduced when making the algorithm generic.  There are 3 obvious choices
of languages to try~\cite{conf/dagstuhl/CzarneckiOST03}: MetaOcaml, 
Template Haskell, and
C++.  Because we wanted our Gaussian Elimination algorithm to ``look like''
the original algorithms in Maple as much as possible while at the same time
enforced as much type-safety as is currently possible, the choice of 
MetaOCaml was quite natural.

\subsection{Algorithm families}

The concept of a \emph{program family}~\cite{journals/tse/Parnas76} is
now an integral
part of software engineering, at least at the design level.  This involves
abstracting all the common parts of a family of (related) programs, and
designing both the common parts and the variation points right at the 
outset.  Knowing the variation points makes a modularization based on
the \emph{design for change}~\cite{journals/cacm/Parnas72a} principle 
that much easier.

This concept of program family can apply to very small programs -- even
programs made up of just one function, as in our case.  This principle
tells us that we should isolate each \emph{aspect}\footnote{as in
separation of concerns, terminology coined by 
Dijkstra~\cite{EWD:EWD447} published as~\cite{EWD:EWD447pub},
the ``other side of the coin'' of modularization according to Parnas.} 
into a separate entity
(a \emph{module} in the case of program families), and then to somehow
paste these aspects together to form a whole.

This is exactly the route we will follow.  However, instead of presenting
the results as a {\it fait accompli}, we will in part show how to 
perform stepwise abstractions (i.e. the inverse steps of stepwise 
refinement~\cite{journals/CACM/Wirth71})
to go from a particular implementation of Gaussian Elimination to a 
generic one.  We believe this to be more instructive.  This presentation
method allows us
to show that, at every step, the resulting source code is quite readable, 
is still
recognizable as Gaussian Elimination, and yet still specializes to the 
original version.

\subsection{Generic and generative programming}

While one can find many different reasons for generic programming
\cite{musser89generic,musser94algorithmoriented,conf/ac/2003gp}, there is one 
fundamentally pragmatic reason:
all decent programmers intensely dislike code duplication.  Code
duplication is sometimes tolerated for two reasons: a poor programming
language which does not have proper abstraction mechanisms to unify
certains language idioms, and efficiency.  These issues are not 
independent: many
abstraction mechanisms (higher-order functions, late-binding, 
ad hoc polymorphism, type classes, etc)
generally incur a run-time cost.  In the case of scientific computation
applications, this run-time cost is frequently unacceptable, at least
in the inner-loop of intensive computations.

Macros and pre-processors were common pseudo-solutions for both of these issues
(and are still heavily used -- many models in physical oceanography are written
in \texttt{Fortran 77} with \texttt{cpp} directives peppered throughout for
configuration choices! \cite{SPOM}).  More recently, generative programming
\cite{Czarnecki} has become an increasingly popular method in scientific
computation - as witnessed by (amongst others) 
Blitz++ \cite{Veldhuizen:1998:ISCOPE}, 
Active Libraries \cite{Veldhuizen:1998:OO98},
GMCL \cite{Czarnecki},
POOMA \cite{POOMA},
BOOST \cite{BOOST},
and ATLAS \cite{ATLAS}.  These have also been applied to product families
\cite{GenerativeProductLines}.  These solutions are mostly
\verb=C++=-centric, which brings problems of its own
\cite{conf/dagstuhl/CzarneckiOST03}.

In a typeful setting, what we want to do is to \emph{expose} the 
configuration parameters of our programs in such a way that the type system
will help rather than hinder, in other words, the type system will 
insure that we are configuring our programs consistently.  In particular, 
we are not interested in being \emph{too} generic: being able to 
use any kind of connector to connect arbitrary components together is
pointless, and will merely result in delaying problems to (at worst)
run-time.

For example, if we have a stateful code type \texttt{code}, the type
of the sequential composition operator of statements \texttt{;} is not
\begin{Verbatim}[commandchars=\\\{\}, 
                 codes={\catcode`$=3\catcode`^=7}]
(;) : code $\times$ code -> code
\end{Verbatim}
but rather
\begin{Verbatim}[commandchars=\\\{\}, 
                 codes={\catcode`$=3\catcode`^=7}]
(;) : (state -> unit) code $\times$ (state -> unit) code 
      -> (state -> unit) code
\end{Verbatim}
where \texttt{state} is a specific representation for a (closed) state type.
This ensures that only statements operating on the same state can be
composed, thereby increasing robustness and safety.

\subsection{Problem}

The Gaussian Elimination algorithm has many different incarnations
(see section~\ref{GEintro}),
which are frequently coded seperately (for efficiency or sometimes
source code clarity reasons).  This is clearly unsatisfactory, as well as
potentially leading to a maintenance nightmare.  On the other hand, 
even when it is possible to use a language's abstraction facility to 
natively express all the design points of GE, this usually has too
high an efficiency cost.

We want to be able to express a generic version of GE which abstracts
out, in the context of symbolic computation, the most common design
points: the domain of the matrix elements ($\mathbb Z$, 
$\mathbb Q$, $\mathbb Z_p$, $\mathbb Q(\alpha)\left[x\right]$, or even
$\mathbb Z_p(\alpha_1,\ldots,\alpha_n)\left[\left[ x_1,\ldots,x_m\right]\right]$
and floating point numbers), the container type used 
(arrays of arrays, flattened arrays, implicit representations, as well 
as variants for storage orientation) for the matrix, the exact 
outputs (matrix, rank, determinant), as well
as some domain-dependent aspects, like whether the elements of the domain
are stored in a normalized manner, or whether there exists a natural (and 
useful) size function on the elements of the domain.  Not only should
this be done in a type-safe manner, the resulting code should also be
recognizable as being Gaussian Elimination.  As well, it should be clear
that the technique could scale to the complete family of algorithms
related to GE, like LU and Cholesky factorization, back-elimination, as
well as more general matrix solving algorithms.

We want our solution to still contain a recognizable Gaussian Elimination
algorithm.  While it is possible to write GE functionally, an imperative
solution is what is generally expected.  The best compromise would be to write
this in something like Haskell's State Monad which, via the do notation,
would \emph{look} right, yet still be functional.  As we did not want to 
add additional complexity by creating some syntactic sugar (via campl4), 
we bit the bullet and dove in headlong into an imperative solution.

\subsection{Contributions}

We show that a number of techniques, 
namely combinators, abstract interpretation
\cite{conf/popl/CousotC77}, state-passing style, and 
dictionary-passing at staging
time can be successfuly combined to achieve our goals of 
writing a generic, efficient, type-safe implementation of the
Gaussian Elimination family of algorithms.  These techniques have
to be very carefully combined, so that whenever a new design point of 
Gaussian Elimination is exposed, 
the generated (source) code does not include new unecessary artifacts, but 
in fact stays
constant when this design point is specialized back to its ``default'' value.

The principal features and contributions of this article are
\begin{itemize}
    \item a thorough analysis of a real situation,
    \item the use of staging for code configuration and clone management,
    \item the analysis of imperative-style code instead of 
		functional, as is more customary,
    \item cleanly documenting and exposing all design-time choices at 
        staging time,
    \item the use of state-passing style (for staging),
    \item and the use of dictionary passing at staging-time.
\end{itemize}

\subsection{Organization}

In section~\ref{GEintro}, we quickly review the Gaussian Elimination algorithm
and examine 3 different variants, to get an idea of the design points
involved.  We then proceed with a more thorough analysis of the design
points that we will tackle in this paper, followed by an analysis of the
relations between these design points.  Section~\ref{staging} quickly
introduces staging, as applied to an abstract Gaussian Elimination 
implementation.  We carefully show how turning this algorithm into a 
sequence of code generators can be done in such a way as to not introduce
any staging artifacts.  We then show how to insert a particular design
choice in to the staged version, and preserve this ``no artifacts''
property.  Section~\ref{techniques} then introduces the various techniques
used in the complete staging of GE.  Only short examples are provided
to illustrate the concepts, but the complete source code is included in an 
appendix.  Related work is reviewed in section~\ref{related}, and 
conclusions are drawn in section~\ref{concl}.  There are 3 appendices, 
one for a fraction-free implementation of GE (in Maple), one for a
feature-rich implementation of GE over algebraic extensions of finite
fields (in Maple), and finally a complete listing of our final implementation
of a staged GE in metaocaml.

\section{Gaussian Elimination}\label{GEintro}

After a quick review of the Gaussian Elimination algorithm, this section
reviews some examples of GE implementations from Maple's library, to get 
an idea of some of the design points involved.  This is followed by a
more thorough analysis of the design points that were encountered, and
which of those will be explicitly treated in this paper.  We draw particular
attention to those design points which are not fully orthogonal to other
design points, as these need special handling to avoid producing
inconsistent or meaningless code.

\subsection{Review of GE algorithm}

Gaussian Elimination is generally understood to be an algorithm which reduces
(square) Matrices to upper-triangular form via a sequence of elementary Matrix
operations, namely row/column interchanges and addition of a row multiple to
another row.  The basic requirements on the the Matrix elements are that they
can be added, multiplied, and zero elements can be recognized.  Many versions
also require that division for non-zero Matrix elements also exists.  The
requirements on the Matrix are that all elements are accessible, and it 
is simple to do elementary row operations (swapping and adding a multiple).

Denote the elements of Matrix $A$ by $a_{ij}$ as usual.
This algorithm proceeds as follows: in the first column, a non-zero element
is chosen, say in row $k$.  If this element is not in the first row, 
exchange row $k$ and the first row.  For every row $j$ below the first one,
add $-\frac{a_{j1}}{a_{11}}$ times the first row to row $j$; this ensures 
that in the resulting row, $a_{j1}=0$.  This is then repeated recursively
on the submatrix obtained by deleting the first row and column.  The
algorithm can be modified to deal with columns with all zeroes, as well as
dealing with non-square matrices.  Details from a numerical analysis point
of view can be found in Chapter 6 of \cite{BurdenFaires}, and for
exact computations in Chapter 9 of \cite{Geddes:1992:ACA}.  The
latter also contains an exposition of the fraction-free version of 
Gaussian Elimination.  One can then show that this version of the 
algorithm works
(ie reduces an $n\times n$ Matrix of rank $n$ to an upper triangular
matrix with a non-zero diagonal) whenever the elements of the Matrix
are from an arbitrary \emph{integral domain}.  There are extensions to
rings which contain zero-divisors as well, but these are beyond the scope
of the current paper.

A frequent variation point is that for reasons of either numerical stability
or coefficient growth, instead of choosing a random row $k$ with a non-zero
element, pivoting is performed by choosing an element that optimizes
a particular metric.  In the case of numerical stability, the largest (in
absolute value) is chosen, whereas for coefficient growth cases, it is 
the smallest (according to some given size metric) that is chosen.  For
numerical stability, full pivoting is also frequently used, although
we will not deal with this here.

\subsection{Some examples}

First, we discuss 3 different implementations of Gaussian Elimination
from Maple's library.  The first one (shown in figure~\ref{gausselim}) 
is the most
straightforward implementation of Gaussian Elimination.  However, it
does include some explicit parametrization, namely the use of
\texttt{Normalizer} and \texttt{Testzero} to allow the use of this
algorithm with any arithmetic domain (in other words one where
the globally defined, polymorphic operations and constants
$+, *, \slash, -$, and $0$ are defined, as well as having a notion of
\texttt{length}).  This implementation  uses Maple's old \texttt{matrix} 
datatype which is in fact implemented via hash tables; while this makes
element access very slow, it does allow very efficient storage of 
sparse matrices, as well as allowing \emph{indexing functions} to be
used for structured matrices.  See the Maple documentation
\cite{Monagan:2001:M7P} for more details on these features.  It is important
to note that this implementation would \textbf{not} be appropriate for 
floating point numbers, as it would be numerically quite unstable.

\begin{figure}
    \begin{small}
\begin{verbatim}
gausselim := proc(A::matrix)
local B,n,m,i,j,c,t,r;
n := rowdim(A); m := coldim(A);
B := map(Normalizer, A); r := 1;
for c to m while r <= n do
    # Search for a provably non-zero pivot element
    i := 0;
    for j from r to n do
        if not Testzero(B[j,c]) then
            if i = 0 then i := j;
            elif length(B[j,c]) < length(B[i,c]) then 
                i := j
            end if; # otherwise B[i,c] is the best
        end if;
    end do;
    if i <> 0 then
        # interchange row i with row r is necessary
        if i <> r then
          for j from c to m do 
            t := B[i,j]; B[i,j] := B[r,j]; B[r,j] := t
          end do
        end if;
        for i from r+1 to n do
          if B[i,c] <> 0 then 
              t := Normalizer(B[i,c]/B[r,c]);
              for j from c+1 to m do 
                B[i,j] := Normalizer(B[i,j]-t*B[r,j]) 
              end do;
              B[i,c] := 0
          end if;
        end do;
        r := r + 1          # go to next row
    end if
end do;              # go to next column
end proc:
\end{verbatim}
\end{small}
\caption{Generic Gaussian Elimination in Maple}
\label{gausselim}
\end{figure}

The second version (see appendix~\ref{FractionFreeElimination}) is
significantly more parametrized: it is implemented via Maple's \texttt{Domains}
package which implements generic programming as originally found in 
Axiom \cite{Jenks:1992:ASC}.  These facilities are modelled on the structures
of classical Algebra like Rings, Fields, Monoids, Groups, etc.  While
the resulting program is very generic, and in Axiom (at least when compiled 
via Aldor \cite{Watt:2002:HCA})
is quite efficient, in an interpreted language like Maple, the results
are unacceptably slow.  In any case, this program implements the fraction-free
version of Gaussian Elimination, which works over any Division Ring.
If the Division Ring is also a Euclidean Domain, then the underlying
Euclidean Norm is used as the notion of \emph{size}, otherwise the 
underlying universally polymorphic \texttt{length} function, which works
for any Maple structure, is used.  Again, one should note that floating
point numbers do not form a Division Ring (as addition is not associative
and multiplication does not distribute over addition).  This implementation
works for any container $A$ (storing the matrix) which allows 2-dimensional
indexing for element retrieval and storage.  Unlike the previous version, 
neither \texttt{Normalizer} nor \texttt{Testzero} are used, which implies
that the underlying rings are assumed to store their elements in 
normalized form, that arithmetic operations return results in normalized
form, and thus zero-testing is straightforward.

The third variant (see Appendix~\ref{modular}) is also quite parametrized,
but along different dimensions: in this case the input domain is fixed
to be an arbitrary algebraic extension (possibly empty) of a finite
field $\mathbb{Z}_p$.  This particular implementation has potentially
multiple outputs: the actual reduced Matrix (always), the rank of the
Matrix (sometimes), and the determinant of the Matrix (sometimes).  It
works for non-square matrices, as well as for augmented matrices, where
one can pass in an integer \texttt{right\_margin} parameter to specify this.
The determinant can only be computed in the case where the principal 
matrix is square.  Elements are not assumed normalized, and for 
efficiency reasons, the normalization procedure used for a pure finite
field is different than the one used for algebraic extensions.  There
is also explicit code for keeping track of the determinant.  This 
version is optimized for the newer \texttt{rtable} data-structure representation
of Maple's newer \texttt{Matrix} constructor, which allows a one-line
expression
\begin{verbatim}
B[i,1..-1], B[row,1..-1] := B[row,1..-1], B[i,1..-1];
\end{verbatim}
to be used for exchanging 2 rows.

For reference, the straightforward implementation of Gaussian Elimination
for matrices of floating point numbers, in OCaml is shown in
figure~\ref{ocaml1}.  It has been coded to resemble the first
Maple version in figure~\ref{gausselim} as much as possible, where 
this makes sense; the most significant difference is that pivoting is
done on the largest number instead of the smallest.

\begin{figure}
    \begin{small}
\begin{verbatim}
let ge_float a =
  let b = copy(a) in
  let n = Array.length b and m = Array.length b.(0) in
  let r = ref 0 and c = ref 0 and i = ref 0 in
  while !c < m && !r < n do
    i := (-1);
    for j = !r to n-1 do
      if not ( b.(j).(!c) = 0.) then 
        if !i = (-1) || 
           abs_float b.(j).(!c) > abs_float( b.(!i).(!c) ) then
             i := j;
    done;
    if !i != (-1) then begin
      if !i <> !r then
        begin
          for j = !c to m-1 do
              let t = b.(!i).(j) in
              b.(!i).(j) <- b.(!r).(j);
              b.(!r).(j) <- t;
          done;
        end;
      for ii = !r+1 to n-1 do
        if not (b.(ii).(!c) = 0.)  then begin
          let t = b.(ii).(!c) /. b.(!r).(!c) in
          for j = !c+1 to m-1 do
              b.(ii).(j) <-  b.(ii).(j) -. t*.b.(!r).(j)
          done;
          b.(ii).(!c) <- 0.;
        end;
      done;
      r := !r + 1;
    end;
    c := !c + 1;
  done ;
  b
  \end{verbatim} 
  \end{small}
  \caption{Floating-point GE in Ocaml}\label{ocaml1}
\end{figure}

\subsection{Design points analysis}\label{designpoints}

A careful review of the rougly 35 implementations of GE we found, as
well as roughly 50 additional implementations of related algorithms, 
a number of implementation patterns became readily apparent.  The main 
variations found were:
\begin{enumerate}
	\item \textbf{Domain}: In which (algebraic) domain do the matrix
		elements belong to.  Some implementations were very specific
		($\mathbb{Z}, \mathbb{Q}, \mathbb{Z_p}, 
		\mathbb{Z}_p\left[\alpha_1,\ldots,\alpha_n\right], 
		\mathbb{Z}\left[x\right]$, $\mathbb{Q}\left(x\right)$, 
		$\mathbb{Q}\left[\alpha\right]$, and floating point numbers for 
		example), while others were generic for elements of a field,
		multivariate polynomials over a field, elements of a formal Ring
		with possibly undecidable zero-equivalence, or elements of a 
		Division Ring.  In the roughly 85 pieces of code we surveyed,
		20 different domains were encountered.
	\item \textbf{Fraction-free}: Whether the Gaussian Elimination
		algorithm is allowed to use unrestricted division, or only
		exact (remainder-free) division.
	\item \textbf{Representation of the matrix}: Whether the matrix
		was represented as an array of arrays, a one-dimensional array,
		a hash table, 
		whether indexing was done in C or Fortran style.  Additionally,
		if a particular representation had a special mechanism for efficient
		row exchanges, this was sometimes used.  Other representations
		were also used.
	\item \textbf{Length measure (for pivoting)}:  For stability reasons
		(whether numerical or coefficient growth), if a domain possesses
		an appropriate length measure, this was sometimes used to choose
		an ``optimal'' pivot for row exchanges.
	\item \textbf{Output choices}:  Whether just the reduced matrix, or
		additional the rank and the determinant were to be returned.
		It is worthwhile noting that in the larger algorithm family,
		routines like \texttt{LinearAlgebra:-LUDecomposition} have
		up to $2^6 + 2^5 + 2^2 = 100$ possible outputs.
	\item \textbf{Normalization and zero-equivalence}: Whether the 
		arithmetic operations of the domain at hand gives results in 
		normalized form, and whether a specialized zero-equivalence 
		routine needs to be used.
\end{enumerate}

\noindent These choices are not completely independent:  
\begin{itemize}
	\item While it is always possible to use full Gaussian Elimination for
		matrices with elements in a Division Ring, the result will in general
		lie in the Field of fractions associated with this Ring.  On the other
		hand, fraction-free GE will always give results in the same domain as
		the input.  There are subtle coefficient growth reasons that come
		into play when deciding, for each domain, the ``best'' version of
		GE to use.
	\item Some domains do not have meaningful length measures.  In particular
		$\mathbb{Z}_p$ is such a field; but by the same token, there is
		no coefficient growth problem with $\mathbb{Z}_p$.  On the other
		hand, even though floating point numbers are of fixed syntactic
		size, numerical stability implies that the absolute value gives
		a meaningful length measure.
	\item While zero-testing is often done via normalization, there are
		times when full normalization is very expensive, and heuristic-driven
		zero-testing can be fruitfully used as a filter for ``easy''
		zero equivalence, before more expensive techniques are used.
\end{itemize}

\noindent A few additional design points were encountered in some cases:
\begin{itemize}
	\item If zero equivalence was known to be undecidable, and a case
		was encountered where a (provably) non-zero pivot could not be
		found, either a warning or an error were issued.
	\item Some implementations had embedded \texttt{userinfo}
		statements, which is Maple lingo for user-controlled directives 
		that give the user information about the execution state of 
		particular algorithms.
	\item Specification that the Matrix was in fact an augmented matrix,
		and that it possessed a natural ``right margin''.  Elimination
		is then performed on the full matrix, but column choices stop
		at this given right margin.  Curiously, while this was frequently
		implemented, this feature was very rarely documented, even though
		this functionality is frequently used in textbooks which 
		outline how Gaussian Elimination can be used on augmented matrices
		to implement solving of linear equations and matrix inversion.
\end{itemize}

\section{Staging Gaussian Elimination}\label{staging}

Using an abstract version of Gaussian Elimination, and assuming that
the reader has some knowledge of staging in MetaOCaml (as expounded in
the other papers in this volume), we show a first na\"{\i}ve staging
of Gaussian Elimination, based on an abstract version of the 
algorithm.

\subsection{Abstract Gaussian Elimination}
Looking at all the Gaussian Elimination algorithms seen so far,
one can extract a staging-free high-level Ocaml version as follows:
\begin{small}
\begin{verbatim}
let ge_high findpivot swap zerobelow a m n =
    let b = copy(a) in
    let r = ref 0 and c = ref 0 in
    while !c < m && !r < n do
        match findpivot b c r n with
        Some i -> begin
            if i <> !r then
                for j = !c to m-1 do
                    swap b i r j;
                done;
            zerobelow b r c n m;
            r := !r + 1;
            end;
        | None -> () ;
        c := !c + 1;
    done ;
    b
\end{verbatim} 
\end{small}

For a version of GE that works with arrays of arrays of floating 
point numbers, the 3 required routines can be given as
\begin{small}
\begin{verbatim}
let findpivot b c r n =
  let i = ref (-1) in begin 
    for j = !r to n-1 do
      if not ( b.(j).(!c) = 0.) then 
        if !i = (-1) || 
          abs_float b.(j).(!c) < abs_float b.(!i).(!c) then
            i := j;
    done;
    if !i== -1 then None else Some !i;
  end;

let swap b i r j =
  let t = b.(i).(j) in begin
    b.(i).(j) <- b.(!r).(j);
    b.(!r).(j) <- t;
  end;

let zerobelow b r c n m = 
  for ii = !r+1 to n-1 do
    if not (b.(ii).(!c) = 0.)  then
      begin
        let t = b.(ii).(!c) /. b.(!r).(!c) in
        for j = !c+1 to m-1 do
          b.(ii).(j) <-  b.(ii).(j) -. t*.b.(!r).(j)
        done;
        b.(ii).(!c) <- 0.;
      end;
    done;
\end{verbatim}
\end{small}

We can recover the previous version \texttt{ge\_float} as
\begin{small}
\begin{verbatim}
let ge_float a = 
    let n = Array.length a and m = Array.length a.(0) in
    ge_high findpivot swap zerobelow a m n
\end{verbatim}
\end{small}

It is however worthwhile to examine closely the type inferred
for \texttt{ge\_high}:
\begin{small}
\begin{verbatim}
# val ge_high : 
  ('a -> int ref -> int ref -> int -> int option) ->
  ('a -> int -> int ref -> int -> 'b) ->
  ('a -> int ref -> int ref -> int -> int -> 'c) -> 
  'a -> int -> int -> 'a = <fun>
\end{verbatim}
\end{small}
\noindent The ``matrix'' \texttt{a} is completely abstracted
out, and we have a fully abstract \verb+'a+ as its type.  Only
when \texttt{findpivot} and friends are instantiated do we get 
this particular type refined.  As well, since the matrix type is
abstract, this also abstracts out the type of elements.
So this abstraction exercise has
already gained us (some) representation independence.

\subsection{First staging}
\noindent  Of course, if one were either greedy for every single
microsecond of efficiency or, as in our case, we are preparing for
a thorough staging of the above routine, we would like
to modify the above code to completely remove all (run-time) traces of 
these abstractions.  This can be done as follows:
\begin{small}
\begin{verbatim}
let ge_gen findpivot swap zerobelow =
    .< fun a m n ->
    let b = copy(a) in
    let r = ref 0 and c = ref 0 in
    while !c < m && !r < n do
        match .~(findpivot_gen .<b>. .<c>. .<r>. .<n>.) with
        Some i -> begin
            if i <> !r then
                for j = !c to m-1 do
                    .~(swap_gen .<b>. .<i>. (lr .<r>.) .<j>.);
                done;
            .~(zerobelow_gen .<b>. .<r>. .<c>. .<n>. .<m>.);
            r := !r + 1;
            end;
        | None -> () ;
        c := !c + 1;
    done ;
    b 
\end{verbatim}
\end{small}
\noindent  While the source code above is not particularly attractive, 
the modifications necessary are certainly quite routine, and
the resulting code is quite readable.  To get
routine \texttt{swap\_gen} from routine \texttt{swap} is also
straightforward:
\begin{small}
\begin{verbatim}
let lr x = .< ! .~ x >.
let swap_gen b i r j =
    .< let b = .~b and i = .~i and j = .~j and r = .~r in
        let t = b.(i).(j) in  begin
        b.(i).(j) <- b.(r).(j);
        b.(r).(j) <- t;
    end >.
\end{verbatim}
\end{small}
\noindent  However, this produces some extra let bindings in the
resulting code (not shown here).  To get back to the original source code,
as intended, the modification is still simple, although the resulting
code (below) now looks quite ugly.  We will assume that these let
bindings are harmless (in other words completely removed by the 
compiler), and use them in our presentation, since we know we could
remove them if needed.
\begin{small}
\begin{verbatim}
let swap_gen2 b i r j =
    .< let t = (.~b).(.~i).(.~j) in  begin
        (.~b).(.~i).(.~j) <- (.~b).(.~r).(.~j);
        (.~b).(.~r).(.~j) <- t;
    end >.
\end{verbatim}
\end{small}
\noindent  Similar modifications to \texttt{zerobelow} and
\texttt{findpivot} are necessary as well.  Once performed, we can
obtain \texttt{ge\_high} verbatim via specialization:
\begin{small}
\begin{verbatim}
let spec_ge = (ge_gen findpivot_gen swap_gen2 zerobelow_gen);
let ge_float2 a = 
    let n = Array.length a and m = Array.length a.(0) in
    let ge = .! spec_ge in
    ge a m n ;
\end{verbatim}
\end{small}
\noindent Close inspection of the program \texttt{spec\_ge} shows that,
up to variable renaming, the resulting code is essentially that 
of \texttt{spec\_float}.
The only real change (and it is not strictly necessary) is the use of
the \texttt{option} type to pass information out from \texttt{findpivot}
instead of using a flag value of $-1$ for the integer reference variable
\texttt{i}.

\subsection{Optimized pivoting}

The first ``design point'' that we will stage is whether we simple use
the first non-zero element we encounter as a pivot, or whether we 
search in the complete column for the ``best'' pivot.  We will thus
assume that we are (potentially) given a function \texttt{better\_pivot}
which, given two values, will return \texttt{true} if the second value
would be a better pivot.  By using an option type, we can also choose
whether we want to use this function.  We parametrize 
\texttt{findpivot\_gen} by this new value, giving:
\begin{small}
\begin{verbatim}
let findpivot_gen better_pivot b c r n =
    let cond i x y = match better_pivot with
    | Some f -> .< !(.~i) = (-1) || .~(f x y) >.
    | None ->   .< !(.~i) = (-1) >.
    in
    .< let i = ref (-1) in begin 
        for j = !(.~r) to .~n-1 do
            if not ( (.~b).(j).(!(.~c)) = 0.) then 
                if .~(cond .<i>. .< (.~b).(j).(!(.~c)) >.
                           .< (.~b).(!i).(! .~c) >. ) then
                    i := j
        done;
        if !i == -1 then None else Some !i;
    end >. ;;
\end{verbatim}
\end{small}
\noindent where we have added a new (local) code-generating
function \texttt{cond} which generates either single condition
\verb+!i = -1+ or splices in the code of our passed-in
\texttt{pivot\_len} function.  Since this is a code-generator,
the body of \texttt{findpivot\_gen} has to be modified accordingly
to pass in the \emph{code} representations necessary for properly
constructing the conditional.

Naturally, we need to make a corresponding change to the instantiation
call.  To reproduce the previous program, this would be done as
\begin{small}
\begin{verbatim}
let spec_ge = 
  let fp = findpivot_gen (Some 
    (fun x y -> .<abs_float .~x < abs_float .~y >. )) in
  ge_gen fp swap_gen2 zerobelow_gen;
\end{verbatim}
\end{small}

\section{Staging techniques}
\label{techniques}

While it is possible to continue staging the code in the na\"{\i}ve
manner described in the previous section, eventually the resulting
source code becomes almost impossible to read.  In other words, this ad hoc
method of staging, as we found out first-hand, does not scale very
well, nor is it always straightforward.
Also, this suggest that 
maintenance would be very challenging -- yet one of the
explicit goals of this exercise to help maintainability.  We need to
find smarter ways to combine and stage code, to simultaneously obtain
readable and efficient programs.  This section presents various
techniques that we have used to successfully stage all of the 
principal design points described in subsection~\ref{designpoints}.

We will use a combination of four techniques: the use of basic
combinators for pieces of code, state passing, abstract interpretation, and 
dictionary passing.  It is important to note that we are staging
(deeply!) imperative code, to obtain imperative code.  Thus we are forced
to develop some techniques which differ from previous work
\cite{TahaThesis,KiselyovTaha} where they were staging purely functional code,
sometimes with the aim of obtaining functional residual code, and sometimes
aiming for imperative residual code.  Of course, some of the techniques
developped below have similar semantic underpinnings as the monadic
approach of \cite{KiselyovTaha}, although this is probably not 
readily apparent when syntactically comparing our results with theirs.

\subsection{Preliminaries}\label{prelim}

Probably the most important point to remember when staging code, especially
in a higher-order language like Ocaml, is that the usual abstraction
technique of using higher-order functions does not quite work as 
expected anymore.  The only object that can be effectively inlined
has to be of \texttt{code} type; in particular it cannot be of
a functional type.  More subtly, we cannot just ``lift'' functions in the
hopes of still being able to apply them before run-time, we instead have to
be careful to deal with (lifted) function bodies.  For example, compare
\begin{small}
\begin{verbatim}
let lift x = .< x >.
let star1 = lift ( * ) in
let f1 x y = .< .~star1 (x+1) (y+3) >. in
f1 2 3 ;;

let star2 x y = .< .~x * .~y >. in
  let f2 x y = star2 .<x+1>. .<y+3>. 
  and f3 x y = let z=x+1 and zz=y+3 in 
      star2 (lift z) (lift zz) 
  and f4 x y = let z=x+1 and zz=y+3 in 
      star2 .<z>. .<zz>. in
[ f2 2 3 ; f3 2 4 ; f4 2 3] ;;
\end{verbatim}
\end{small}
\noindent which gives as results
\begin{small}
\begin{verbatim}
# - : ('a, int) code =
.<(((* compiled code omitted (as id: x) *)) 
    (2 + 1) (3 + 3))>.
# - : ('a, int) code list =
[.<((2 + 1) * (3 + 3))>.;
 .<((* compiled code omitted (as id: x) *) *
    (* compiled code omitted (as id: x) *))>.;
 .<(3 * 6)>.]
\end{verbatim}
\end{small}

All of these are operationally equivalent!  However, only for
\texttt{f4} do we get what we were really after.
Careful reading of \cite{TahaThesis} explains all of these, but
they are still a bit disconcerting to the novice MetaOCaml programmer.
While the difference between \texttt{f2} and \texttt{f4} is easy to
explain in terms of binding-time, the difference between \texttt{f1}
and \texttt{f4} takes more getting used to for the seasoned functional
programmer, who is used to passing around functions.  It is very 
important to pass around code instead - which is the main difference
between \texttt{star1} and \texttt{star2}: the first is completely
closed code which can only applied at run-time, while the latter can
still be applied at stage 1.  This may be obvious to meta programming
gurus, but is not emphasized enough in practical introductions to
staging in MetaOCaml.

\subsection{Combinators}

It is very natural to try to develop \emph{code combinators}, to attempt
to do for code what worked so well for parsers \cite{ParserCombinators}.
While this works to a certain extent, the next technique in part 
destroys the applicability of pure code combinators to many cases of interest,
as we are here mainly dealing with combining imperative program fragments.

While we experimented with many pure code combinators, in the end 
the only one we actually used was a simple lift of sequencing:

\begin{small}
\begin{verbatim}
let seq a b = .< begin .~a ; .~b end >.
\end{verbatim}
\end{small}

\noindent It is worthwhile noting that the begin/end are needed, otherwise
MetaOCaml complains that this is syntactically invalid.
Another useful combinator (used in development) was composition:

\begin{small}
\begin{verbatim}
let compose a b = fun x -> .< .~a ( .~b x ) >.
\end{verbatim}
\end{small}

\noindent While it is possible to create many more combinators, like
\begin{small}
\begin{verbatim}
let _if_ cond body = .<if .~cond then .~body >. ;;
\end{verbatim}
\end{small}
\noindent in actual use we have found that this ends up making the code
much less readable than staging larger pieces of code.  But the reader
should remember that our aims are quite different than that of other authors,
so these conclusions a priori might only apply to our particular setting.

\subsection{State passing}

The first really new technique (in the context of staging) is that
of \emph{state passing style}.  This is taken simultaneously from two
different sources: the state-transformer paradigm of denotational
semantics for programming languages \cite{DenSem}, the StateMonad
and ST monads of Haskell, and Sumii's work
\cite{sumii01hybrid}
on hybrid online-offline partial evaluation, where a state-passing style
was used over a continuation-passing style, to great effect.

While in these general paradigms, the state space is usually 
open\footnote{in other words is potentially infinite}, we want much tighter
control on the state.  So we use an explicit record type to encode our
state configuration, with the finite
set of state variables that our complete program will use.  We can then
rely on the type-checker to make sure that this is the only state we use.
For example, for our needs we define
\begin{small}
\begin{verbatim}
type 'a state = 
    {b:'a array array; r:int ref; c:int ref; m:int; n:int}
\end{verbatim}
\end{small}
\noindent which then forces us to modify all our generators
to refer to this state configuration explicitly, and we have to add record
dereferences as appropriate.
\begin{small}
\begin{verbatim}
let swap_gen s i j =
    .< let s = .~s and i= .~i and j= .~j in
       let t = s.b.(i).(j) in  begin
        s.b.(i).(j) <- s.b.(!(s.r)).(j);
        s.b.(!(s.r)).(j) <- t;
    end >.
\end{verbatim}
\end{small}
\noindent  While this does make some code a bit harder to read, this
has (essentially) no run-time cost, and definitely cuts down on the 
number of parameters
needed for each generator.  The main generator then needs to be modified
to create this state configuration
\begin{small}
\begin{verbatim}
.< fun a -> 
   let s = {b=a; r=ref 0; c=ref 0; m=Array.length a.(0); 
         n=Array.length a} in
  ... >.
\end{verbatim}
\end{small}

Note that it is possible to remove even this (potential) overhead of
record selection, by defining our state configuration to contain
code generators instead of values.  See subsection~\ref{dict} for
an example of this.

\subsection{Abstract interpretation}

Programmers coming from systems that permit unbounded manipulation of 
white-box code (Scheme, Maple, Mathematica, etc) need to remember that
white-box code cannot exist
in MetaOCaml.  Thus any information about a piece of code that is necessary to
be re-used later must be explicitly exposed, since such information cannot be
seen once it is encapsulated in a code type.

Suppose we know a priori that the generic code contains 
\verb+ if !i = -1 then+, but that sometimes we would like to modify
this to\\
{\small\verb+if !i = -1 ||+\\
\verb+abs_float b.(j).(!c) > abs_float( b.(!i).(!c) ) then+},
if we had white-box code, this could be done by actually finding that code,
and explicitly modifying it.  In metaocaml, this technique is not available.
There are reasons for this (like voiding equational reasoning principles
on computations inside brackets \cite{TahaThesis}, as this essentially
reduces the brackets to syntactic questions \cite{SICP}).
Instead, one must explicitly insert a code generator\\
\verb+if .~(cond_gen .<i>. etc) then+ which will conditionally generate
the correct code.  The question then becomes, how do we indeed 
conditionally generate the correct code for each particular instance?

We use a simple case of the general technique of \emph{abstract interpretation} 
\cite{conf/popl/CousotC77} to solve this problem; see \cite{KiselyovTaha} for a
different use of the same technique.  The main idea is to take the fully
abstract code type, and to add some additional information to it, 
information which can be used at staging time to generate more precise
code.  This additional information is an approximation of the complete
information contained in the white-box code.
In our case, this translates to knowing whether our domain contains
the functions necessary to optimize the choice of a pivot.  Thus we 
define a type
\begin{small}
\begin{verbatim}
type pivoting_code_option =
  (('a,'b) code -> ('a,'b) code -> (('a,bool) code)) option;
\end{verbatim}
\end{small}
\noindent which says that we \textit{might} have a ``smaller than'' 
boolean code generator available.  This is used with a conditional
code generator
\begin{small}
\begin{verbatim}
let orcond_gen main_cond = function
    | Some alt_cond -> .< .~main_cond || .~alt_cond >.
    | None -> main_cond ;;
\end{verbatim}
\end{small}
\noindent to generate the required code depending on the situation.
A more complex use of the same technique is when a domain requires a 
normalization step; in that case, we have to insert the normalization
code into the abstract code for mapping over our generic two-dimensional
container.  We end up using code that looks like
\begin{small}
\begin{verbatim}
let mdapply1 mapper g c1 = match g with
    | Some f -> .< .~(mapper f c1) >.
    | None   -> c1 ;

.~(mdapply1 mapper normalizerf .<a>. )
\end{verbatim}
\end{small}
where \texttt{mapper} is a code generator for mapping a function \texttt{f}
(also given as code) over the code of an abstract container.  Also 
see section~\ref{concl} for a more substantial application of abstract
interpretation.

\subsection{Dictionaries}\label{dict}

While it is possible to use just the techniques of the previous sections
to stage a fairly generic GE, the resulting source code would be hideous.  In
particular, most generators would have between 15 to 20 parameters, 
since a typical domain has 10 different components, and a generic 
container has 5.  Instead, it makes a lot more sense to gather these
parameters together into a type.

To be more specific, let us look at domains.  Since they are essentially
Division Rings, we know that they must contain the constants zero and one, 
as well as the operations of addition, subtraction, multiplication
and (a restricted kind of) division.  It may optionally come with
a ``smaller than'' relation, a normalizer, and a zero-tester.  For
simplicity, we can also assume that we are given an explicit $-1$ value
as well.  So we could define a type corresponding to a domain over
some arbitrary base set:
\begin{small}
\begin{verbatim}
type 'b domain = {zero:'b; one:'b; minusone:'b;
    plus:'b -> 'b -> 'b; times:'b -> 'b -> 'b;
    minus:'b -> 'b -> 'b; div:'b -> 'b ->'b;
    smaller_than:('b -> 'b -> bool) option;
    normalizer:('b -> 'b ) option;
} ;;
\end{verbatim}
\end{small}
\noindent It should be remarked that what we are essentially 
building there is a type for a (static) dictionary for domains.
This is akin to (simple) type classes in Haskell, or indeed of
defining a module in OCaml over an abstract type.

If we lift this domain up one stage, then via partial application 
and inlining, we could hope to completely eliminate the overhead
of this level of genericity.  But doing things this way would run
into the problems already outlined in subsection~\ref{prelim}.
While we would get a program which is operationally correct, and which
Ocaml's underlying optimizer \emph{might} resolve to what we want,
we certainly have no guarantees of that.  To be quite sure of 
this, we need to provide not the values and functions that
correspond to a Division Ring, but rather the code (and code
generators) for each of these.  In other words, the type we really
need is
\begin{small}
\begin{verbatim}
type ('a,'b) domain = {zero:('a,'b) code; one:('a,'b) code;
  minusone:('a,'b) code;
  plus:('a,'b) code -> ('a,'b) code -> ('a,'b) code; 
  times:('a,'b) code -> ('a,'b) code -> ('a,'b) code;
  minus:('a,'b) code -> ('a,'b) code -> ('a,'b) code; 
  div:('a,'b) code -> ('a,'b) code ->('a,'b) code;
  smaller_than:(('a,'b) code -> ('a,'b) code -> 
               (('a,bool) code)) option;
  normalizerf:(('a,'b -> 'b) code) option;
  normalizerg:(('a,'b) code -> ('a,'b) code) option;
} ;;
\end{verbatim}
\end{small}
\noindent  We need two different versions of the normalizer function
because we might use the normalizer in two different ways: either
in an imperative setting, like in the body of a \texttt{for} loop, 
or in a functional setting, like as an argument to \texttt{Array.map}.

For example, one can define the domain of integers $\mathbb{Z}$ as
\begin{small}
\begin{verbatim}
let dom_int = { zero = .< 0 >. ; one = .< 1 >. ;
   minusone = .< -1 >. ;
   plus = (fun x y -> .<.~x + .~y>.) ;
   times = (fun x y -> .<.~x * .~y>.) ;
   minus = (fun x y -> .<.~x - .~y>.) ;
   div = (fun x y -> .<.~x / .~y>.) ;
   smaller_than = Some(fun x y -> .< abs .~x < abs .~y >.) ;
   normalizerf = None ;
   normalizerg = None }
\end{verbatim}
\end{small}
\noindent where we have been careful to define code-generating
functions rather than code-of-functions.  We have also completely
eschewed the use of \texttt{lift}, to ensure we get complete
inlining.  This might be ultimately be unecessary, but we would have 
to trust the MetaOCaml \texttt{run} function \texttt{.!} to 
systematically do this for us, which\footnote{for theoretically sound 
reasons!} we cannot rely on.

Using similar techniques, we can also completely abstract out
the matrix container as follows:
\begin{small}
\begin{verbatim}
type ('a,'c,'d) container2d = {
  get: ('a,'c) code -> ('a,int) code -> 
       ('a,int) code -> ('a,'d) code ;
  set: ('a,'c) code -> ('a,int) code -> 
       ('a,int) code -> ('a,'d) code -> ('a,unit) code ;
  dim1: ('a,'c) code -> ('a,int) code;
  dim2: ('a,'c) code -> ('a,int) code;
  mapper: ('a, 'd->'d) code -> ('a,'c) code -> ('a,'c) code
} 
\end{verbatim}
\end{small}
\noindent where \texttt{'a} is the extra type variable required by the
abstract code type, \texttt{'c} is the abstract container type, and
\texttt{'d} is the abstract element type.  Even though from 
\texttt{get} and \texttt{set}, we could derive the code for \texttt{mapper},
letting the user give an explicit \texttt{mapper} functions allows for
extra efficiency.  This could certainly be considered to be an
additional design point.

It is worthwhile noting that these 2 types are completely independent, 
and only when are they actually used together will the type system
unify the 3rd type variable of \texttt{container2d}
with the 2nd type variable from \texttt{domain}.  This is quite pleasing,
as it reassures us that we have an appropriate separation of concerns
between the domain of elements and the abstract container.

These 2 types have values that are fully available at staging time.  
Thus any overhead that could be associated with these abstractions 
is in fact completely eliminated, and the generated code is free
of these abstractions.  In other words, we were able to use staging
instead of defunctionalisation \cite{defunc} 
to remove a complete layer of abstraction.

\section{Staging Gaussian Elimination, continued}

The techniques of the previous section allow us to take care of all
of the major design points that were outlined in subsection~\ref{designpoints}.
For example, by using a combination of abstract interpretation and a
code combinator, we can track which information is needed in the final
code, allowing the final code generator for the swap routine to look
like
\begin{small}\begin{verbatim}
let swapr_gen dom contr track s i j =
  let swap =
  .< let b = (.~s).b and r = !((.~s).r)
      and i= .~i and j= .~j in
     let t = .~(contr.get .<b>. .<i>. .<j>. ) in  begin
       .~(seq
           (contr.set .<b>. .<i>. .<j>. 
		             (contr.get .<b>. .<r>. .<j>. ))
           (contr.set .<b>. .<r>. .<j>. .<t>. ) );
  end >. in
  let tds = 
     .< (.~s).detsign := 
	   .~(dom.times .<!((.~s).detsign)>. dom.minusone) >. in
  match track with
  | TrackNothing -> swap
  | TrackDet     -> seq swap tds
\end{verbatim}\end{small}
\noindent where we see the use of the abstract container \texttt{cont},
the abstract domain \texttt{dom}, the tracking information
\texttt{track}, and the combinator \texttt{seq}.

The one area that was slightly less satisfying was in staging the
choice of outputs.  While any given instantiation of GE will produce
a specific output, the family of algorithms exposes a choice.  Since
typing in MetaOCaml is done once, this choice persists in the
\emph{type}
of the final program, even though it should be readily apparent that
this choice does not in fact persist in the actual generated code.
More precisely, we use a type
\begin{small}\begin{verbatim}
type ('b,'c) outputs = 
      Matrix of 'b
    | MatrixRank of 'b * int
    | MatrixDet of 'b * 'c
    | MatrixRankDet of 'b * int * 'c
\end{verbatim}\end{small}
\noindent for the output of GE.  Clearly any generated code will only
use one of these choices, but this will not be apparent in the 
type of the generated code.

In the end, the type of our function \texttt{specializer} is
\begin{small}\begin{verbatim}
val specializer : ('a, 'b) domain ->
  ('a, 'c, 'b) container2d -> fracfree:bool ->
  outputs:outchoice -> ('a, 'c -> ('c, 'b) outputs) code
\end{verbatim}\end{small}
\noindent which says that it outputs code for a 
function of type \verb+'c -> ('c, 'b) outputs+ with the 
constraints that \verb+'c+ be a container type and \verb+'b+ be
a domain type.  A complete listing, with a specialization, is
given in Appendix~ref{finalcode}.

\section{Related work}\label{related}

There is now quite an established line of work in scientific (numerical)
programming of using generative techniques (mainly in C++, see 
\cite{Czarnecki,Veldhuizen:1998:ISCOPE,ATLAS} and the work they cite).
However, their worries are somewhat different than ours: they mainly
aim for efficiency - they are able to generate code which is
usually faster than the handwritten Fortran programs they replace.
While they do not overlook maintainability issues, that is often a 
secondary concern.
This drive for efficiency usually means worrying about many low-level
details (cache sizes, cache misses, etc), whereas we are much more 
interested in writing very generic
programs which can hopefully still be quite efficient.
While clearly very successful, with FFTW \cite{FFTW98} being the most famous 
this work cannot claim to be type-safe.  In the case of FFTW, the work
of Kiselyov, Swadi, and Taha \cite{KiselyovTaha} is probably the first
that can claim type safety.

Our work is much more closely related to the work on generic programming
coming out of Symbolic Computation and Computer Algebra
\cite{musser89generic,Jenks:1992:ASC}.  Computer Algebra systems, starting
with ScratchPad \cite{ScratchPad}, on through Axiom \cite{Jenks:1992:ASC} and
Aldor \cite{Watt:2002:HCA} worked very hard at modelling the structures
encountered in mathematics (Groups, Rings, Fields, Modules, etc)
through abstract data types in various programming languages.  This
allowed them to code many algorithms very generically, allowing
extremely general mathematical structures to be defined and computed with.
Unfortunately, this very genericity came at an extremely high efficiency
cost; part of this is due to the fact that many mathematical structures
(with Fields and Matricies being good examples) are not inductive data
types.  In fact, any reasonable type system for properly reflecting 
mathematics requires dependent types (as was already known to the 
AUTOMATH people in 1968 \cite{Bruijn83}, but has been 
re-discovered many times).  In Maple itself, the \emph{Domains}
package (based on an older package called Gauss \cite{Gruntz:1994:IG})
successfully implements this strategy, but at a whopping factor of
30 efficiency cost over base Maple, which is already a factor of 60
slower than comparable C code.  This is why generative techiques must
be used to (at least) regain that last factor of 30.  What we show here
is that (in the context of MetaOCaml), this indeed works.

Others have used partial evaluation techniques to solve the
``configuration'' problem, closely related to what we have done here.
In particular, the work of Consel and his colleagues
\cite{lemeur-al:hosc03,Marlet-Thibault-Consel:jase99} is relevant.  They use
a declarative language to declare the specialization points of 
components; the components are written in a mainstream programming language
(here C), without the need for annotations.  While very pragmatic, this
inherits all of the defects of the host language - and in this case
this means that type-safety is completely lost.  Additionally, it is
clear that such an approach could not be used to express nearly as many
design points as we have been able to express in quite compact code.
Nevertheless, some of the underlying ideas of this work is very close
to ours.

Lastly, it should be noted that the programs we end up writing are very
reminiscent of hand-written language interpreters in cogen style
\cite{conf/pepm/BondorfD94,conf/icfp/Thiemann96}.  In hindsight,
this is not really surprising, as this is the style that MetaOCaml
forces upon us, by not having white-box code. Thus we have to be writing
code generating functions.

\section{Conclusion}\label{concl}

We have shown how to combine several techniques 
namely combinators, abstract interpretation, state-passing style, and 
dictionary passing (at staging time) can be 
carefully and successfuly combined to achieve our goals of 
writing a generic, efficient, type-safe implementation of the
Gaussian Elimination family of algorithms.  Furthermore, 
the generated code does not include unecessary artifacts introduced
by generalizing the code over several design points.

It should be easy to see this these techniques generalize immediately
to the larger family of algorithms containing LU decomposition, Cholesky 
factorization, forward as well as backward elimination, and of course
linear system solvers.

\subsection{Design points revisited}

Reviewing the design points listed in subsection~\ref{designpoints},
we have shown how to abstract out domains, whether the algorithm should
be fraction-free or not, matrix representations, the use of a size
measure (when available) for pivoting, output choices, and the
potential need for normalization.

On the other hand, we have not shown how to deal with domains where
zero-testing is undecidable, nor in fact on how to use a special
zero-testing routine when it is given; both of these are straightforward
using the techniques shown here.

Something else which has not been adequately addressed is whether
a particular domain requires that a fraction-free algorithm be used
or not -- in effect whether the input domain is in fact a Field or 
just a Division Ring.  We could simply add a new boolean field 
\texttt{is\_field} to our domain type, but this is not very elegant.
It would be much more pleasant if the compiler had some Theorem 
Proving capabilities, and we could instead let the compiler decide
this property.

Although we showed how to use a domain's length function, we are
hard-coding the \texttt{smaller\_than} function.  Akin to the
decision to use a fraction-free algorithm, it would be much better
to let the compiler somehow prove that we are either in a numerical
domain that suffers from instability issues, or in a symbolic domain
which suffers from coefficient growth issues, and conclude how
the length function should be used.

While we have shown how to deal with output choices, our approach
clearly does not scale: given something like the current implementations for 
LU Decomposition in Maple, we would have to use an algebraic type
with 100 constructors.  This is clearly not a realistic solution.
However, in these cases, a more sophisticated type system could
perhaps deal with the issue: the output type is quite simple at
stage 0.  Thus if there could be \emph{staged types} as well as
staged code, this problem could be easily handled.  Another intriguing
possibility would be to use GADTs.  Naturally, this could also be
handled via dependent types, but in this instance, this seems like
overkill.

Lastly, the container type could be abstracted out even more.  Our
algorithm right now is only really well-suited to dense rectangular
matrices.  Any kind of structured matrix, especially a sparse matrix,
could benefit from a differently structured traversal strategy. 
Simple issues like faster methods for row exchange would be quite
straightforward to handle, but handling sparsity is more challenging.
We believe that the techniques we show here are sufficient to handle
this as well, but would require a more significant restructuring of
the code generators.  Note that we have also completely ignored 
some issues
of frequent concern in implementations of Gaussian Elimination for
large matrices of floating point numbers, like full pivoting, 
blocking, etc.  Full pivoting would be quite straightforward to 
implement, whereas blocking would be more challenging.

\subsection{Future work}

We intend to deal with structured matrices next, as well as actually
implementing larger portions of the algorithm family centered around
Gaussian Elimination.  We then hope to continue generalizing this work --
as it is known that Gr\"{o}bner bases degenerate to Gaussian Elimination
in the case of linear systems, it would be interesting to see if we could
reproduce this via staging as well.  This is definitely considerably
more difficult, as linearity is a \emph{property} of the input system.
Dealing with the complete family of algorithms related to Gaussian 
Elimination, and linear algebra algorithms in general, should be
relatively straightforward, and we intend to complete this work to
ensure this is indeed the case.

We also want to pursue more aggressively the issue of abstract interpretation.
One item that intrigues us is how straightforward it is in our setting
to generalize the code shown in section 2 of~\cite{KiselyovTaha}:
\begin{small}
\begin{verbatim}
type ('b,'c) abstract = Elem of 'b | Any of 'c 
type ('a,'b) monoid = {one: 'b; op:'b->'b->'b;
    op_gen:('a,'b) code->('a,'b) code->('a,'b) code}

let concretize = function 
    | Elem x -> .<x>.
    | Any  x -> x
and generalize x = Elem x

let special_op dom a y = 
    if a=dom.one then 
        y 
    else 
        Any (dom.op_gen .<a>. .< .~(concretize y) >.)

let gen_op dom = fun x y ->
    match (x,y) with
    | (Elem a, Elem b) -> Elem (dom.op a b)
    | (Elem a, y)      -> special_op dom a y
    | (x, Elem b)      -> special_op dom b x
    | (x, y)           -> Any (dom.op_gen 
      .< .~(concretize x) >. .< .~(concretize y) >.)

\end{verbatim}
\end{small}
This code generates an abstract operator for any monoid.
We need to specify the (actual) identity for the monoid, as 
well as the monoid operation and a code generator for
the monoid information.  Then we use abstract interpretation
to evaluate the monoid operations as early as possible.  Note that
here we are able to evaluate any monoid operation, and not just
operations involving the identity.  
We can easily specialize this to the case covered in
\cite{KiselyovTaha} via
\begin{small}
\begin{verbatim}
let ( ** ) x y = gen {
    one = 1 ; op = (fun x y -> x * y);
    op_gen = fun x y -> .< .~x * .~y >. } x y 
let rec power n x = 
    if n=0 then generalize 1 else x ** (power (n-1) x)
let power3 = 
    .< fun x -> .~(concretize (power 3 (Any .<x>.))) >.
\end{verbatim}
\end{small}
and we get the expected result of
\begin{small}
\begin{verbatim}
val power3 = .< fun x -> x*x*x >.
\end{verbatim}
\end{small}
Better still, the version of power above, apart from the use of
\texttt{generalize}, looks very much like the power code that one would
na\"{\i}vely write (modifying it to do binary powering is straightforward).  
The implementation above assumes a commutative monoid, which
is frequently the case (and is the case here), but it can easily be
generalized to the non-commutative case.  On the other hand, the 
above version is easily generalized to do normalization of 
multivariate monomials over an arbitrary base monoid.

While we have used records here to encode all our abstract types,
we really should have used Ocaml's modules, as that is the proper way
to encode dictionaries in Ocaml.  If we then use Functors, for both
input and output configuration~\cite{KiselyovPersonal}, we can make
our implementation even clearer, as well as offering a solution to the
problem of configurable output types.

Lastly, we want to investigate the additional power that using a
combination of monadic and continuation passing style may bring.  It
would appear that staging certain complex aspects (like determinant
tracking) can be made considerably simpler by using this style.

\subsection{Acknowledgments}

This work started with a visit by the author to Walid Taha while the latter
was still at Yale.  Thanks to his gentle prodding, this paper finally 
got written.  This paper benefited from several useful conversations
with Bill Farmer on the topics of genericity, generative software, and
compilers as theorem provers.  Thanks go to David Linder for helping
the author catalogue all the implementations of Gaussian-Elimination
related algorithms in Maple, and Fr\'{e}d\'{e}ric Chyzak for drawing
the authors' attention (years ago) to this issue.  Not least, Oleg 
Kiselyov made some very very fruitful comments on a draft of this paper, for
which author is grateful.

\bibliographystyle{elsart-num}
\bibliography{ge}

\appendix
%Appendix A
\section{Maple code for \texttt{FractionFreeElimination}}
\label{FractionFreeElimination}
\begin{small}
\begin{verbatim}
FractionFreeElimination := proc(C,M,A,m,n) 
  local d,i,j,k,s,t,size;
  size := `if`(hasCategory(C,EuclideanDomain), 
               C[EuclideanNorm], length);
  s := 1; d := C[1];
  for k to m-1 do
      for i from k to m while A[i,k] = C[0] do end do;
      if i > m then return C[0] fi;
      for j from i+1 to m do
          if A[j,k] <> C[0] and 
              size(A[j,k]) < size(A[i,k]) then 
              i := j 
          end if
      end do;
      #  Pivot is A[i,k], interchange if necessary
      if i <> k then
        s := - s;
        for j from k to n do 
          t := A[k,j]; A[k,j] := A[i,j]; A[i,j] := t;
        end do;
      end if;
      #  Fraction-free elimination
      for i from k+1 to m do
          for j from k+1 to n do 
              t := C[`-`]( C[`*`](A[i,j],A[k,k]), 
                           C[`*`](A[k,j],A[i,k]));
              A[i,j] := C[Div](t, d)
          end do;
      end do;
     d := A[k,k]
  end do;
  s
end proc:
\end{verbatim}
\end{small}
%Appendix B
\section{Maple code for \texttt{`mod/rtable/Gausselim`}}
\label{modular}
\begin{small}
\begin{verbatim}
`mod/rtable/Gausselim` := proc(A::Matrix,rank::{integer,name},
                               det::{integer,name})
  local    B, d, nrows, ncols, right_margin, i,j,
      col, row, s, t, p, Q;

  Q := type( A, 'Matrix( rational )' );
  if not Q and not type( A, 'Matrix( ratpoly( algnum ) )' ) then
      error "first argument must be a matrix over a finite field"
  end if;
  ( nrows, ncols ) := LinearAlgebra:-Dimensions( A );

  p := args[-1];
  right_margin := ncols;
  if type( args[-2], 'integer' ) then
      right_margin := args[-2]
  end if;

  if nargs > 3 and type( det, 'name' ) 
               and nrows <> right_margin then
    error "cannot compute determinant of a non-square matrix"
  end if;

  if Q then
      B := map( s -> s mod p,  A );
  else
      B := map( s -> 'Normal'( s ) mod p, A );
  end if;

  row := 1; # initialise the row
  d := 1; # initialise the determinant

  for col to right_margin while row <= nrows do
      # Find the first nonzero entry in this column
      for i from row to nrows while B[ i, col ] = 0 do
          # empty
      end do;
      if i > nrows then
          # found a zero column so determinant is zero
          d := 0;
          next
      end if;
      if i <> row then
          # interchange row i with row row
          B[ i, 1..-1 ], B[ row, 1..-1 ] := 
              B[ row, 1..-1 ], B[ i, 1 .. -1 ];
          d := -d mod p; # keep track of determinant
      end if;

      if nargs = 4 then
          d := 'Normal'( d * B[ row, col ] ) mod p
      end if;

      # Make all the nonzero entries below this one 
      # are equal to 1
      if Q then
          s := 1/B[ row, col ] mod p;
          for i from row + 1 to nrows do
              # do i-th row
              if B[ i, col ] = 0 then
                  next
              end if;
              t := s * B[ i, col ] mod p;
              for j from col + 1 to ncols do
                  B[ i, j ] := B[ i, j ] - 
                               t * B[ row, j ] mod p
              end do;
              B[ i, col ] := 0
          end do
      else
          s := 'Normal'( 1 / B[ row, col ] ) mod p;
          for i from row + 1 to nrows do
              if B[ i, col ] = 0 then
                  next
              end if;
              t := 'Normal'( s * B[ i, col ] ) mod p;
              for j from col + 1 to ncols do
                B[i,j] := 'Normal'( B[ i, j ] - 
                                    t * B[ row, j ] ) mod p
              end do;
              B[ i, col ] := 0
          end do
      end if;

      row := 1 + row # go to next row
  end do; # go to next column

  if nargs > 2 and type( rank, 'name' ) then
      rank := row-1
  end if;
  if nargs > 3 and type( det, 'name' ) then
      det := d
  end if;
  B
end proc:
\end{verbatim}
\end{small}
%Appendix C
\section{Final MetaOCaml code}
\label{finalcode}
\begin{small}
\begin{verbatim}
type ('a,'c,'d) container2d = {
  get: ('a,'c) code -> ('a,int) code -> 
       ('a,int) code -> ('a,'d) code ;
  set: ('a,'c) code -> ('a,int) code -> 
       ('a,int) code -> ('a,'d) code -> ('a,unit) code;
  dim1: ('a,'c) code -> ('a,int) code;
  dim2: ('a,'c) code -> ('a,int) code;
  mapper: ('a, 'd->'d) code -> ('a,'c) code -> 
          ('a,'c) code;
  copy: ('a, 'c) code -> ('a,'c) code
} ;;

type ('a,'c,'d) state = {b:('a,'c) code; 
  r:('a,int ref) code; c:('a,int ref) code; 
  m:('a,int) code; n:('a,int) code;
  det:('a,'d ref) code; detsign:('a,'d ref) code};;

type ('a,'b) domain = {
  zero:('a,'b) code; one:('a,'b) code;
  minusone:('a,'b) code;
  plus:('a,'b) code -> ('a,'b) code -> ('a,'b) code; 
  times:('a,'b) code -> ('a,'b) code -> ('a,'b) code;
  minus:('a,'b) code -> ('a,'b) code -> ('a,'b) code; 
  div:('a,'b) code -> ('a,'b) code ->('a,'b) code;
  smaller_than:(('a,'b) code -> ('a,'b) code -> 
                ('a,bool) code) option;
  normalizerf:(('a,'b -> 'b) code ) option;
  normalizerg:(('a,'b) code -> ('a,'b) code ) option;
} ;;

type ('b,'c) outputs = 
    Matrix of 'b
  | MatrixRank of 'b * int
  | MatrixDet of 'b * 'c
  | MatrixRankDet of 'b * int * 'c ;;

type outchoice = JustMatrix | Rank | Det | RankDet;;
type dettrack = TrackNothing | TrackDet ;;

type ge_choices = 
  {fracfree:bool; track:dettrack; outputs:outchoice} ;;

let orcond_gen main_cond = function
  | Some alt_cond -> .< .~main_cond || .~alt_cond >.
  | None -> main_cond ;;

let seq a b = .< begin .~a ; .~b end >. ;;

let sapply2 g c1 c2 = match g with
  | Some f -> Some (f c1 c2)
  | None   -> None ;;

let dapply1 g c1 = match g with
  | Some f -> f c1
  | None   -> c1 ;;

let mdapply1 mapper g c1 = match g with
  | Some f -> .< .~(mapper f c1) >.
  | None   -> c1 ;;

let choose_output dom s = function
  | JustMatrix -> .< Matrix (.~s.b) >.
  | Rank       -> .< MatrixRank((.~s.b), ! .~(s.r)) >.
  | Det        -> .< MatrixDet((.~s.b), 
     .~(dom.times .<! .~(s.det)>. .<! .~(s.detsign)>. ))>.
  | RankDet    -> .<MatrixRankDet((.~s.b), 
    ! .~(s.r), .~(dom.times .<! .~(s.det)>.
    .< ! .~(s.detsign) >. )) >.

let ge_state_gen dom contr findpivot swap zerobelow choice=
  let main_loop s = 
  .< while !(.~s.c) < .~s.m && !(.~s.r) < .~s.n do
    begin
    match .~(findpivot s) with
    Some i -> begin
        if i <> !(.~s.r) then
            for j = !(.~s.c) to .~s.m-1 do
                .~(swap s .<i>. .<j>.);
            done;
        .~(zerobelow s) ;
        (.~s.r) := !(.~s.r) + 1;
        end;
    | None -> .~(match choice.track with
        | TrackNothing -> .< () >. ;
        | TrackDet     -> .< .~s.detsign := .~dom.zero >.;
        ) ;
    end;
    .~s.c := !(.~s.c) + 1;
  done >. in

  .< fun a -> 
  let b= .~(mdapply1 contr.mapper dom.normalizerf 
                     (contr.copy .<a>. )) and
      r=ref 0 and c=ref 0 and
      m= .~(contr.dim1 .<a>. ) and 
      n= .~(contr.dim2 .<a>. ) and
      det=ref .~(dom.one) and detsign=ref .~(dom.one) in
  .~(let st = {b= .<b>.; r= .<r>.; c= .<c>.; m= .<m>.; 
               n= .<n>.; det= .<det>.; 
               detsign= .<detsign>.} in
    seq
      (main_loop st )
      (choose_output dom st choice.outputs ) ) 
  >. ;;

let swapr_gen dom contr track s i j =
  let swap =
  .< let b = .~(s.b) and r = !(.~s.r) and i= .~i 
     and j= .~j in
      let t = .~(contr.get .<b>. .<i>. .<j>. ) in  begin
        .~(seq
            (contr.set .<b>. .<i>. .<j>. 
               (contr.get .<b>. .<r>. .<j>. ))
            (contr.set .<b>. .<r>. .<j>. .<t>. ) );
  end >. in
  let tds = .< .~s.detsign := 
        .~(dom.times .<! .~s.detsign>. dom.minusone) >. in
  match track with
  | TrackNothing -> swap
  | TrackDet     -> seq swap tds ;;

let findpivot_gen dom contr orcond_gen = fun s ->
  .< let i = ref (-1) in begin 
    for j = ! .~s.r to .~s.n-1 do
      if not ( .~(contr.get .< .~s.b>. .<j>. 
                 .<!( .~s.c)>. ) = .~(dom.zero)) then 
        if (.~(orcond_gen .< !i = (-1) >.
          (sapply2 dom.smaller_than 
            .< .~(contr.get .< .~s.b>. .<j>. .<! .~s.c>. )>.
            .< .~(contr.get .< .~s.b>. .<!i>. .<! .~s.c>. )>.
            ))) then
          i := j;
      done;
      if !i == -1 then None else Some !i;
  end >. ;;

let zerobelow_gen dom contr choice s =
  let inner_loop = 
  if not choice.fracfree then fun i s ->
    .< let t = .~(dom.div 
      (contr.get .<.~s.b>. .< .~i>. .<!(.~s.c)>. )
      (contr.get .<.~s.b>. .<!(.~s.r)>. .<!(.~s.c)>. ) ) in
    for j = !(.~s.c)+1 to .~s.m-1 do
      .~(contr.set .<.~s.b>. .<.~i>. .<j>. 
          (dapply1 dom.normalizerg 
          (dom.minus (contr.get .<.~s.b>. .<.~i>. .<j>.)
          (dom.times .<t>. 
              (contr.get .<.~s.b>. .<!(.~s.r)>. .<j>. )))))
    done; >.
  else fun i s -> 
    .< for j = !(.~s.c)+1 to .~s.m-1 do
      let t = .~(dapply1 dom.normalizerg (dom.minus 
        (dom.times 
          (contr.get .<.~s.b>. .<.~i>. .<j>. )
          (contr.get .<.~s.b>. .<! .~s.r>. .<! .~s.c>. ))
        (dom.times 
          (contr.get .<.~s.b>. .<!(.~s.r)>. .<j>. )
          (contr.get .<.~s.b>. .<.~i>. .<! .~s.r>. )))) in
      .~(contr.set .<.~s.b>. .<.~i>. .<j>. 
            (dom.div .<t>. .<! .~s.det>. ))
    done >. in
  let outer_loop =
  .< for i = !(.~s.r)+1 to .~s.n-1 do
    if not ( .~(contr.get .<.~s.b>. .<i>. .<! .~s.c>. ) 
             = .~dom.zero) then
      begin
        .~(inner_loop .<i>. s );
        .~(contr.set .<.~s.b>. .<i>. .<! .~s.c>. dom.zero)
      end;
  done >. in
  match choice.track with
  | TrackNothing -> outer_loop
  | TrackDet     -> seq outer_loop
    (if choice.fracfree then
      .< .~s.det := 
        .~(contr.get .<.~s.b>. .<! .~s.r>. .<! .~s.c>. )>.
    else
      .< .~s.det := .~(dom.times .<! .~s.det>. 
          (contr.get .<.~s.b>.
              .<! .~s.r>. .<! .~s.c>. ))>. );;

let specializer dom container ~fracfree ~outputs =
  let t = (if fracfree || outputs == Det 
                       || outputs == RankDet then 
      TrackDet else TrackNothing ) in
  let choice = 
      {fracfree=fracfree; track=t; outputs=outputs} in
  let fp = findpivot_gen dom container orcond_gen in
  let swap = swapr_gen dom container choice.track in
  let zb = zerobelow_gen dom container choice in
  ge_state_gen dom container fp swap zb choice;;

let dom_float = { 
   zero = .< 0. >. ; 
   one = .< 1. >. ;
   minusone = .< -1. >. ;
   plus = (fun x y -> .<.~x +. .~y>.) ;
   times = (fun x y -> .<.~x *. .~y>.) ;
   minus = (fun x y -> .<.~x -. .~y>.) ;
   div = (fun x y -> .<.~x /. .~y>.) ;
   smaller_than = 
       Some(fun x y -> .<abs_float .~x < abs_float .~y >.);
   normalizerf = None ;
   normalizerg = None };;

let array_container = {
  get = (fun x n m -> .< (.~x).(.~n).(.~m) >. );
  set = (fun x n m y -> .< (.~x).(.~n).(.~m) <- .~y >. );
  dim2 = (fun x -> .< Array.length .~x >.) ;
  dim1 = (fun x -> .< Array.length (.~x).(0) >. ) ;
  mapper = (fun f a -> .< Array.map 
      (fun x -> Array.map .~f x) .~a >.);
  copy = (fun a -> .<Array.map (fun x -> Array.copy x) 
                       (Array.copy .~a) >. )
};;

let spec_ge_float = 
    specializer dom_float array_container 
                ~fracfree:false ~outputs:RankDet ;;

let ge_float3 = .! spec_ge_float ;;
\end{verbatim}
\end{small}
\end{document}