### 2007-11-21 tagless/talk.tex
Start to work on the APLAS 2007 talk (nothing to see here yet)

\documentclass[ucs]{beamer}
\usepackage{helvet}
\usepackage[utf8x]{inputenc}
\usepackage{microtype}
\usepackage{calc}
%\usepackage{tikz}

\usepackage{textcomp}
\makeatletter
\uc@dclc{8249}{default}{\textlangle}
\uc@dclc{8250}{default}{\textrangle}
\makeatother

\usepackage[compact]{fancyvrb1}
\DefineVerbatimEnvironment{code}{Verbatim}{commandchars=\\\{\}}

\def\>{\leavevmode\rlap{\color{blue}$\blacktriangleright$}\ \ }
\newcommand\slide[2]{#2<#1>}

\setbeamertemplate{section in toc}{{\leavevmode\llap{$\blacktriangleright$ }\bfseries\inserttocsection}\par}
\setbeamertemplate{section in toc shaded}{{\color{black}\inserttocsection}\par}
\setbeamertemplate{subsection in toc shaded}[default][50]

\AtBeginSection[]
{
  \begin{frame}<beamer>
    \frametitle{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

\begin{document}

\newcommand{\authorstack}[3]{\begin{footnotesize}\begin{tabular}[t]{@{}c@{}}\begin{normalsize}#1\end{normalsize}\\[\jot]#2\\\texttt{#3}\end{tabular}\end{footnotesize}\ignorespaces}
\newsavebox\carette \sbox\carette{\authorstack{Jacques Carette}{McMaster University}{carette@mcmaster.ca}}
\newsavebox\oleg    \sbox\oleg   {\authorstack{Oleg Kiselyov}{FNMOC}{oleg@pobox.com}}
\newsavebox\ccshan  \sbox\ccshan {\authorstack{Chung-chieh Shan}{Rutgers University}{ccshan@rutgers.edu}}

\setbeamertemplate{sidebar right}{}

\title{Finally tagless, partially evaluated}
\subtitle{Tagless staged interpreters for simpler typed languages}
\author[Jacques Carette and Oleg Kiselyov and Chung-chieh Shan]{\usebox\carette \and \usebox\oleg \and \usebox\ccshan}
\date{30 November 2007}

\begin{frame}
    \titlepage
\end{frame}

\setbeamertemplate{background}{%
    \raisebox{0pt}[\paperheight][0pt]
        {\includegraphics[angle=90,width=\paperwidth+1pt]{White_pearl_necklace.jpg}}%
    \llap{\href{http://www.flickr.com/photos/28481088@N00/160781390/}
               {\scriptsize\color{gray}tanakawho on flickr\kern1pt}}}
\begin{frame}
\end{frame}
\setbeamertemplate{background}[default]

\setbeamertemplate{sidebar right}{\vfill\llap{\usebeamerfont{framesubtitle}\insertpagenumber\hskip\jot}\vskip\jot}

\begin{frame}
\end{frame}

\end{document}