### 2007-11-25 tagless/talk.tex
Introduce the object language

\pagenumbering{arabic}
\documentclass[ucs,professionalfont]{beamer}
\usepackage{concmath}
\usepackage{helvet}
%\renewcommand{\rmdefault}{cmss}
%\usepackage[utf8x]{inputenc}
\usepackage[override]{cmtt}
\usepackage{mdwtab}
\usepackage{prooftree1} \proofrulebaseline=1.5ex \proofdotseparation=.75ex
\usepackage{microtype}
\usepackage{calc}

\usepackage{tikz}
\usepgflibrary{shapes}
\usepgflibrary{arrows}

\usepackage{textcomp}
\makeatletter
\uc@dclc{8249}{default}{\textlangle}
\uc@dclc{8250}{default}{\textrangle}
\makeatother

\usepackage[compact]{fancyvrb1}
\DefineVerbatimEnvironment{code}{Verbatim}{commandchars=\\\{\}}

\def\>{\leavevmode\rlap{\color{blue}$\blacktriangleright$}\ \ }
\newcommand\slide[2]{#2<#1>}

%\newcommand{\ZZ}{\mathbold{Z}}
%\newcommand{\BB}{\mathbold{B}}
\newcommand{\ZZ}{\textrm{int}}
\newcommand{\BB}{\textrm{bool}}
\newcommand{\fun}[1]{\mathopen{\lambda\mathord{#1}.\;}}
\newcommand{\fix}[1]{\mathopen{\mathop{\textrm{fix}}\mathord{#1}.\;}}
\newcommand{\cond}[2]{\mathopen{\textrm{if}}\;#1\mathrel{\textrm{then}}#2\mathrel{\textrm{else}}}

\setbeamertemplate{section in toc}{{\leavevmode\llap{$\blacktriangleright$ }\bfseries\inserttocsection}\par}
\setbeamertemplate{section in toc shaded}{{\color{black}\inserttocsection}\par}
\setbeamertemplate{subsection in toc shaded}[default][50]

\AtBeginSection[]
{
  \begin{frame}<beamer>
    \frametitle{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

% To get the page numbers to stay constant on a frame [dpt]
\mode<presentation>
{
  \setbeamertemplate{sidebar left}{\thispdfpagelabel{\insertframenumber}}
}

\title{Finally tagless, partially evaluated}
\subtitle{Tagless staged interpreters for simpler typed languages}
\author{Jacques Carette and Oleg Kiselyov and Chung-chieh Shan}

\begin{document}

\newcommand{\authorstack}[3]{\begin{footnotesize}\begin{tabular}[t]{@{}c@{}}\begin{normalsize}#1\end{normalsize}\\[\jot]#2\\\texttt{#3}\end{tabular}\end{footnotesize}\ignorespaces}
\newsavebox\carette \sbox\carette{\authorstack{Jacques Carette}{McMaster University}{carette@mcmaster.ca}}
\newsavebox\oleg    \sbox\oleg   {\authorstack{Oleg Kiselyov}{FNMOC}{oleg@pobox.com}}
\newsavebox\ccshan  \sbox\ccshan {\authorstack{Chung-chieh Shan}{Rutgers University}{ccshan@rutgers.edu}}

\setbeamertemplate{sidebar right}{}

\author[Jacques Carette and Oleg Kiselyov and Chung-chieh Shan]{\usebox\carette \and \usebox\oleg \and \usebox\ccshan}
\date{APLAS\\30 November 2007}

\begin{frame}
    \titlepage
\end{frame}

\setbeamertemplate{background}{%
    \raisebox{0pt}[\paperheight][0pt]
        {\includegraphics[width=\paperwidth+1pt,trim=0 100 0 0]{White_pearl_necklace.jpg}}%
    \llap{\href{http://www.flickr.com/photos/28481088@N00/160781390/}
               {\scriptsize\color{gray}tanakawho on flickr\kern1pt}}}
\begin{frame}
\end{frame}
\setbeamertemplate{background}[default]

\setbeamertemplate{sidebar right}{\vfill\llap{\usebeamerfont{framesubtitle}\insertframenumber\hskip\jot}\vskip\jot}

\tikzstyle{lang}=[ellipse,draw,anchor=west,text depth=3pt,text height=9pt]
\tikzstyle{every picture}=[>=angle 60]

\setbeamertemplate{background}{%
    \raisebox{0pt}[\paperheight][0pt]
        {\includegraphics[width=\paperwidth,trim=0 35 0 180,clip]{FC0836213122.jpg}}}
\begin{frame}<1>[t,label=everywhere]{There's interpretation everywhere}
    A fold on an inductive data type is an interpreter of a domain-specific
    language.

    \bigskip
    \begin{tikzpicture}
        \draw node [lang] (lang0) {contract};
        \foreach \label/\prev/\cur in {grammar/0/1,music/1/2,$\lambda$-term/2/3}
            \draw (lang\prev.east)+(.3,0) node [lang] (lang\cur) {\label};
        \draw (lang3.east)+(.2,0) node [anchor=west] {\dots};
    \onslide<2>
        \draw (lang0)+(-110:2cm)   node (terp01) {value};
        \draw (lang0)+( -85:2.5cm) node (terp02) {schedule};
        \draw (lang1)+(-110:2cm)   node (terp11) {parse};
        \draw (lang1)+( -85:2.5cm) node (terp12) {pretty-print};
        \draw (lang2)+(-110:2cm)   node (terp21) {typeset};
        \draw (lang2)+( -85:2.5cm) node (terp22) {perform};
        \draw (lang3)+(-110:2cm)   node (terp31) {evaluate};
        \draw (lang3)+( -85:2.5cm) node (terp32) {compile};
        \foreach \lang in {0,...,3} {
            \draw (lang\lang)+(-65:1.5cm) node {\dots};
            \foreach \terp in {1,...,2}
                \draw [->] (lang\lang) -- (terp\lang\terp);
        }
    \onslide<3>
        \draw (lang3)+(-148:4.5cm)  node [inner xsep=0pt] (terp31) {evaluate};
        \draw (lang3)+(-133:4.25cm) node [inner xsep=0pt] (terp32) {compile};
        \draw (lang3)+(-113:4cm)    node [inner xsep=0pt] (terp33) {specialize};
        \draw (lang3)+( -93:4.4cm)  node [inner xsep=0pt] (terp34) {CPS transform};
        \draw (lang3)+( -73:4cm)    node [inner xsep=0pt] (terp35) {pretty-print};
        \draw (lang3)+( -58:3cm)    node {\dots};
        \foreach \terp in {1,...,5} \draw [->] (lang3) -- (terp3\terp);
    \end{tikzpicture}

    \only<2>{The same language can be interpreted in many useful ways.}%
    \only<3>{We focus on the $\lambda$-calculus as an example.}
\end{frame}
\setbeamertemplate{background}[default]
\againframe<2->[t]{everywhere}

\begin{frame}{Simple type preservation}
    \def\bool{\only<7-8>{\ \alert{bool}}}
    \begin{tikzpicture}
        \draw node [lang] (lang)
          {\begin{overprint}[12em]
               \onslide<1>\makebox[12em]{\alert{Typed} source language}
               \onslide<2>\makebox[12em]{$3 \le 4$}
               \onslide<3-7>\makebox[12em]{\texttt{LEQ (LIT 3, LIT 4):\makebox[4em][l]{\bool\ term}}}
               \onslide<8>\makebox[12em]{\texttt{\alert{leq} (\alert{lit} 3, \alert{lit} 4):\makebox[4em][l]{\bool}}}
               \onslide<9>\makebox[12em]{Simply typed $\lambda$-calculus}
           \end{overprint}};
        \draw [->] (lang) --
            node [above,midway,anchor=east]
              {\only<1>{\alert{Typed} metalanguage\hspace*{2em}}%
               \only<2-3>{evaluate}%
               \only<9>{\begin{tabular}{@{}r@{\qquad}}
                        Haskell constructor instances\\
                        or ML functors
                        \end{tabular}}}
            node [above,midway,anchor=east,text width=2.5in,inner xsep=0pt]
                {\only<-3,9>{\color{bg}}\ttfamily
                \alt<8->{\only<8>{\alert<8>{lit:\ int -> int}\\
                lit i = i\\[1ex]
                \alert<8>{leq:\ int * int -> bool}\\
                leq (i,j) = i <= j}}
                {\only<3-7>{\alert<7>{eval:\ \only<7>{'a }term -> \only<7>{'a }value}\\
                eval (LIT i) = INT i\\
                eval (LEQ (m,n)) =\\
                \quad \alert<5>{match (eval m, eval n) with\\
                \quad (INT i, INT j) ->} BOOL (i <= j)}}}
            +(-100:4cm)
            node [anchor=north,text depth=0pt,text height=9pt]
              {\llap{\onslide<3-7>{\texttt{BOOL }}%
                     \onslide<3-8>{\texttt{true:}}%
                     \only<1>{\rlap{\alert{Typed} target languages}}%
                     \only<2>{\rlap{\enspace\textrm{true}}}%
                     \only<9>{\rlap{evaluate}}}%
               \makebox[3em][l]{\texttt{\bool\only<3-7>{\ value}}}};
        \draw [->] (lang) -- +(-80:3.5cm)
            node [anchor=north west,inner xsep=0pt] {\only<9>{\hspace*{-16pt}compile}};
        \draw [->] (lang) -- +(-65:3cm)
            node [anchor=north west,inner xsep=0pt] {\only<9>{specialize}};
        \draw [->] (lang) -- +(-50:2.5cm)
            node [anchor=north west,inner xsep=0pt] {\only<9>{\hspace*{-3pt}CPS transform}};
        \draw      (lang)    +(-35:2cm) node {\dots};
    \end{tikzpicture}

    \bigskip

    \strut
    \only<-4>{It should be obvious in the metalanguage that interpreting
    a well-typed source term yields a well-typed target term.}%
    \only<5-6>{The term should be \alert{\alt<5>{well-typed}{closed}}, so
    \alert{\alt<5>{pattern matching}{environment lookup}} in the metalanguage should always \textbf{obviously} succeed.}%
    \only<7>{Previous solutions use (and motivate) fancier types:\\
    generalized abstract data types (GADT) and dependent types.}%
    \only<8>{Our simple solution is to be \textbf{finally tagless:}\\
    replace term constructors by cogen functions.}%
    \only<9>{The term accommodates \textbf{multiple interpretations}\\
    by abstracting over the cogen functions and their types.}%
    \strut
\end{frame}

\section{The object language}
\subsection{As a constructor class in Haskell}

\begin{frame}[t,fragile]{The object language\only<3-8>{ as a constructor class}\only<9-16>{ as a functor signature}}
\setbeamercovered{transparent}
\begin{overprint}
\onslide<1-2>
\medskip
\begin{proofrules}
    \[ \[ [x:t_1] \proofoverdots e:t_2 \] \justifies \fun{x}e:t_1\to t_2 \]
    \[ \[ [f:t_1\to t_2] \proofoverdots e:t_1\to t_2 \] \justifies \fix{f}e:t_1\to t_2 \]
    \[ e_1:t_1\to t_2 \quad e_2:t_1 \justifies e_1 e_2: t_2 \]
    \[ \text{$n$ is an integer} \justifies n:\ZZ \]
    \uncover<1>{\[ \text{$b$ is a boolean} \justifies b:\BB \]}
    \[ e_1:\ZZ \quad e_2:\ZZ \justifies e_1+e_2:\ZZ \]
    \uncover<1>{\[ e_1:\ZZ \quad e_2:\ZZ \justifies e_1 \times e_2:\ZZ \]}
    \uncover<1>{\[ e_1:\ZZ \quad e_2:\ZZ \justifies e_1 \le e_2:\BB \]}
    \[ e:\BB \quad e_1:t \quad e_2:t \justifies \cond{e}{e_1}{e_2}:t \]
\end{proofrules}
\onslide<3-8>
\begin{semiverbatim}
class Symantics \alert<3>{repr} where
     \alert<4,5>{int }\alert<5>{:: Int -> repr Int}
     \alert<4,6>{lam }\alert<6>{:: (repr a -> repr b) -> repr (a -> b)}
     \alert<4,6>{fix }\alert<6>{:: (repr a -> repr a) -> repr a}
     \alert<4,7>{app }\alert<7>{:: repr (a -> b) -> repr a -> repr b}
     \alert<4  >{add }:: repr Int -> repr Int -> repr Int
     \alert<4,8>{if_ }\alert<8>{:: repr Bool -> repr a -> repr a -> repr a}
\end{semiverbatim}
\onslide<9-16>
\begin{semiverbatim}
module type Symantics = sig type \alert<9>{\smash{('c,'v)}\,repr}
 val \alert<10,11>{int}\alert<11>{:\,int\,->\,('c,int)\,repr}
 val \alert<10,12>{lam}\alert<12>{:\,(('c,'a)\,repr\,->\,('c,'b)\,repr)\,->\,('c,'a->'b)\,repr}
 val \alert<10,12>{fix}\alert<12>{:\,('x\,->\,'x)\,->\,(('c,'a->'b)\,repr as 'x)}
 val \alert<10,13>{app}\alert<13>{:\,('c,'a\,->\,'b)\,repr\,->\,('c,'a)\,repr\,->\,('c,'b)\,repr}
 val \alert<10   >{add}:\,('c,int)\,repr\,->\,('c,int)\,repr\,->\,('c,int)\,repr
 val \alert<10,14>{if_}\alert<14>{:\,('c,bool)\,repr\,->\,(unit\,->\,'x)\,->\,(unit\,->\,'x)}
     \alert<10,14>{   }\alert<14>{ \,         \,    \,->\,(('c,'a)\,repr as 'x)}
end
\end{semiverbatim}
\end{overprint}

\onslide<1->
\begin{tabular}{@{}Mll@{}}
\only<5-8,11-14>{\text{\textcolor{structure.fg}{Object term}}}%
\only<16>{\rlap{\textcolor{structure.fg}{ML higher-order functor}}}
& \only<5-8,11-15>{\textcolor{structure.fg}{\only<-14>{\llap{\tikz\draw[->,semithick](0,0)--(1.5,0)node[anchor=mid,inner sep=0pt]{\vphantom{O}};\:}}\only<3-8>{Haskell term}\only<9-15>{ML higher-order functor}}} \\
\multicolumn{2}{@{}l@{}}{\only<16>{\alert{\texttt{functor (S:Symantics) -> struct open S}}}} \\
\only<16>{\rlap{\alert{\texttt{ let term () =}}}}%
\visible<1-14>{\uncover<1-2,5-8,11-14>{\alert<6,12>{\fun{x} \fix{f} \fun{n}}}}
& \ttfamily
  \only<5-8>{\alert<6>{lam (\textbackslash x -> fix (\textbackslash f -> lam (\textbackslash n ->}}%
  \only<11-16>{\alert<12>{lam\,(fun x\,-> fix\,(fun f\,-> lam\,(fun n\,->}} \\
\visible<1-14>{\uncover<1-2,5-8,11-14>{\alert<8,14>{\cond{\textcolor{black}{n\le\alert<5,11>{0}}}{\textcolor{black}{1}}}}}
& \ttfamily
  \only<5-8>{\alert<8>{if\textunderscore} \alert<8>(leq n (\alert<5>{int 0})\alert<8>) \alert<8>(int 1\alert<8>)}%
  \only<11-16>{\alert<14>{if\textunderscore} \alert<14>(leq n (\alert<11>{int 0})\alert<14>) \alert<14>{(fun\,()\,->}\,int 1\alert<14>)} \\
\visible<1-14>{\uncover<1-2,5-8,11-14>{x\times \alert<7,13>{f(n-1)}}}
& \ttfamily
  \only<5-8>{\alert<8>(mul x (\alert<7>{app f (add n (int (-1)))})\alert<8>))))}%
  \only<11-16>{\alert<14>{(fun\,()\,->}\,mul x\,(\alert<13>{app f\,(add n\,(int\,(-1)\!)\!)}\!)\!\alert<14>)\!)\!)\!)} \\
\multicolumn{2}{@{}l@{}}{\only<16>{\alert{\texttt{end:\ functor (S:Symantics) -> sig}}}} \\
\only<16>{\rlap{\alert{\texttt{ val term:\ unit\,->}}}}%
\visible<1-14>{\uncover<1-2,5-8,11-14>{\mathord:\;\ZZ\to\ZZ}}
& \ttfamily
  \only<5-8>{::\ Symantics repr => repr (Int -> Int)}%
  \only<11-16>{('c,int->int)\,S.repr} \\
\multicolumn{2}{@{}l@{}}{\only<16>{\alert{\texttt{end}}}}
\end{tabular}
\end{frame}

\subsection{As a functor signature in ML}
\section{Tagless interpretation}
\subsection{Evaluation}
\subsection{Compilation}
\section{Type-indexed types}
\subsection{Partial evaluation}
\subsection{CPS transformation}

\begin{frame}{Conclusion}
    Initial type-checking can be done

    Preserves sharing
\end{frame}

\end{document}