### 2008-06-21 SCP2008/paper/reply-to-reviews.txt
draft reply to the reviewers

	We are very grateful to the reviewers for their comments and
so helpful suggestions. We have implemented them all. Below is the
detailed response (omitting suggestions about spelling and grammar,
which we have corrected as recommended).

Reviewer #1:

The summary of our work given by the reviewer is profound; we couldn't
have said better ourselves.

Reviewer #2:

> Two larger changes I suggest are restructuring section 4 and performing
> stronger evaluation of the work.
We have implemented them, as described below.

> Restructuring section 4
> The subsections of section 4 are rather random and disjoint. There
> needs to be a more cohesive plan about what is going on in this
> section. Better yet a concrete example that gets built upon in each
> subsection and motivates the subsection's functionality. Sections 4.1,
> 4.3, 4.5, 4.6 seem to be about defining modules for particular
> aspects. Part of Section 4.2 seems superfluous from the comment
> above. The "DOMAIN" module seems superfluous too: you never need to do
> real operations on values, but only generate code for operations on
> values. Hence "DOMAINL" is what you need and "DOMAIN" is not
> needed. Also I don't think "CONTAINER2D" is ever seen again after
> being introduced in Section 4.2 -- this seems troubling.

We added a paragraph at the end of the container section clarifying
that the `current' container is available to the algorithmic aspects,
under the name of |C|. We describe in more detail the `multi-stage'
nature of instantiating the generic algorithm, pointing out which
module is a part of which.

> Section 4.4 is about programming in OCaml and how to maintain the
> algorithm's state.  Section 4.7 is about putting it all together.

> From my understanding of your work here is an idea to make your paper more
> precise and concise.
>  - Move Section 4.4 into Section 3 because that section seems to deal with
>    low-level programming issues.
Moved.

> - In the intro of Section 4 make a link to the list in Section 2 and say your
>   going to pick some of those aspects and explain how they are implemented -
>   one per subsection.
Done

> - Delete the first half of 4.2.
> - Combine 4.1 with the second half of 4.2, dropping the use of "DOMAIN" and
>   just do everything in terms of "DOMAINL".

It seems DOMAIN is a nice and simple abstraction to build upon later;
it also corresponds to our introductory example in Sec 2. We have
completely re-written Sec 4.2 (which is now Sec 5.2), but decided to 
leave DOMAIN in.

> - Sections 4.3, 4.5, 4.6 then become Sections 4.2, 4.3, 4.4.
> - Section 4.7 could be the last subsection of 4 
> or a new whole section with more detail.

This plan has being followed through. Hopefully the introduction to
(now) Sec 5 describes the cohesive plan of the section.

> Strengthening evaluation
> I understand that program generators are complex programs, that
> explaining them can be difficult and that providing concise and
> meaningful code snippets about them can also be hard. Nevertheless a
> good paper on programming program generators needs to do these
> things. At the end of your paper I thought you had laid down all the
> pieces of the machinery well, but I didn't have much of a sense of how
> the machine runs. So I kind of have to take your word that everything
> works, especially since the code on p24-5 is not explained in detail
> and the appendix code is not readable (by itself it's not very
> convincing: how can I tell this code demonstrates something good?). In
> general, and this would strengthen the merit of the paper, code
> snippets need to be explained better - too often I see a new
> identifier in the code that is not explained.

The comparison of the generated code with the hand-written code
hopefully tells now that the generated code does do the right
thing. We have expanded many explanations for the code snippets,
explicitly relating them to the hand-written sample code -- at the
expense of the length of the paper.

> Showing and explaining the code for the unstaged algorithm might help
> and would give a point of reference to compare the generated code
> with. Explaining the generated code could also be helpful, but maybe
> too hard. Maybe it would be possible to focus on one or more small
> areas of the algorithm and show: this is what the area looks like in
> unstaged program, this is the part of the code generator responsible
> for producing code for this area, and this is the area in the
> generated code. Something like this would give the reader a greater
> understanding of how logic flows through your generation process and
> how your functor technique is ultimately in control of fine-grained
> details of code building.

We have done exactly that. We added Sec 4 with the hand-written
unstaged code, and refer to it when discussing various aspects,
especially the determinant aspect.

> p23, "off-shoring": give relevant citation
Done. Incidentally, Walid Taha spells offshoring without the hyphen,
so we have adjusted the spelling.

> p2, line 1, For an introduction paragraph this could be clearer and stronger.
> What are the "aspects of linear algebra"? How are they addressed and by what?
> (I know they are explained later on.) Why is "may" emphasized?
We re-wrote the paragraph.

> p2, line 6, LU decomposition is a particular case of Gaussian Elimination, not
> vice-versa
We have made that clear all throughout the text (noting that it is
quite common, especially in the naming of library functions, to treat
GE and LU as quasi-synonyms)

> p2, line 10, this paper doesn't not discuss
> hardware-architecture-specific optimization, in fact native code
> generation seems beyond the scope of this paper, I think all you want
> to say is that generative programming allows the programmer to express
> domain-specific optimizations
Yes, we did

> p4, "The design space", this decomposition of the design space is very
> interesting and an important result, but it would be good if this
> section was more compact. E.g., maybe present the aspects in their
> groups straight up instead of two lists.
That is true. However, as this is important, it needs to
be explained in detail, so that all the relevant information is recorded,
in writing. We do consider this identification of aspects as one of
the contributions.

> p10, Figure 1, I don't follow the intuition behind naming the
> operation "retN"; I don't see how the operation is similar to ordinary
> monadic "ret", or what the "N" means; wouldn't a name like "letM" be
> better?
We have added the explanation of the name (on p11 now). Also we chose
retN for consistency with Swadi et al, where RetN was first
introduced.

> p10, Figure 1, showing the type of "retN" and maybe some other values
> would be helpful
Done for ret, retN and bind.

> p10, the type variable 'p and its constraint: could you explain the
> problem with classifiers in more detail? if the type was "type ('s,
> 'w, 'v) monad" what's wrong with having types like "('a state, 'w,
> ('a, 'v) code) monad"?
There is more explanation at the end of p11 and beginning of
p12. There is nothing wrong, in principle, with the type 
"('a state, 'w,  ('a, 'v) code) monad" other than the fact the types
become bigger and bigger; furthermore, only a couple of components of
the type are really needed when giving a type signature.

> p14, line 4, don't you want to assert that "ZpMake" is a "DOMAIN" as was done
> for "IntegerDomain"?
Actually no: we omitted the annotation on purpose, to demonstrate that
we can do it. This is further explained in the text (the end of p18).

> p15, shouldn't the code here appear as the "code type" implementation
> of an "abstract code" module?
> p15, what are "Logic" and "Idx" used for?
> p16, what is the point of the "S" module? why can't "DOMAINL" be
> parameterized by "T" (containing "rep") directly?
> p16, Section 4.2, none of the abstract code combinators were used in
> the second half of this section: was the first half necessary? what is
> the relationship between these two halves that warrants them being in
> the same subsection?  p16, Section 4.2, I understand the purposes of
> abstracting code representation.  Additionally it adds to the
> genericity of the code generator, which is good, but doesn't incur any
> abstraction overhead, which is also good. However for the purposes of
> this paper it is perhaps an abstraction too far: the paper never needs
> to generate anything other than MetaOCaml code and it increases the
> complexity of code snippets. Mentioning that you do abstract code
> representation in the real implementation but elide it for the paper
> might make the paper more succinct and clearer.

The section on abstract code and lifted domain has been completely
re-written, hopefully the above questions are all answered, in
particular about the meaning of Logic and Idx and the merits of
abstracting over interpretations.

> p17, line 4, where does the identifier "rep" come from in this code?
> p17, line -2, what accompanying code?
> p18, line 4, what is the "decl" function?
> p19, showing the types of these functions would be helpful
> p19, line -3, what is the "DETERMINANT" module? it hasn't been introduced yet
> p20, what are "cmonad" and "omonad"?

Sec 5.2 has been completely re-written; hopefully it now answers all
these questions,

> p17, line 4, what is the design decision for making "Dom" a definition
> inside "CONTAINER2D" instead of a parameter of "CONTAINER2D"? For
> readers not familiar with the OCaml module system it might be helpful
> to provide a brief introduction somewhere.
> p17, line -12, now in the module implementation "Dom" is made functor
> parameter, why? and why can you just replace "rep" with "code"?
We added a new paragraph to explain this, the last but one paragraph 
of Sec 5.3.

> p18, although most readers are probably familiar with simple LU
> decomposition, a brief overview that shows how determinant and rank
> are computed in the algorithm with an unstaged code example showing
> the state needed for tracking would be helpful to demonstrate the
> problems stated here
We added a new section, Sec 4.

> p20, shouldn't there be an "answer" method in the state objects?
This is now covered by proper definition of cmonad/omonad.

> p19, line -2, why do you need to track the determinant for integer
> matrices
We explained that in the new Sec 4.

> p20, what is "classif" used for? initially I thought it meant "class
> if", but I just worked out it means "classifier", so this relates to
> an earlier comment, so I think it is important you explain in detail
> why you need to put the classifier in this object.
That is now explained on p11.

> p20, line 5, "e.g., a piece of code...": the type of "lstate" depends on the
> implementation and neither of the implementations shown set "lstate" to "v
> ref".
AbstractDet contains a tdet, which is defined to be v ref
one line above which, by the open statement right above it, it Dom.v ref.

> p21, keep the names of the sign and magnitude(?) state consistent
> across "decl", "upd_sign" and "fin"; also keep the code to retrieve
> the sign and magnitude state consistent across "upd_sign" and "fin"
We have done that.

> p20, what do "acc", "get", "set" do?
Changed.

> p21, what is "wmatrix"? shouldn't this be an instance of "CONTAINER2D"?
It is now defined. We explained its differences from the CONTAINER2D
(the very end of p29).

> p21, "(...,res) monad" what does "..." mean?
We made it more precise.

> p21, the "unitL" in "decl" is superfluous, right? "mo_extend" will
> already return a monadic unit result
Actually no. mo_extend has the result type ('a,unit) monad whereas
unitL has the result type ('a,unit) cmonad, which corresponds to
('a,('c,unit) code) monad. In other words, the former describes an
action that generates no code whereas the latter is an action that
generates code (such as an assignment operation) that has the type
unit. However, we did mean that decl is an action to be executed
solely for its side effects, and it was used as such in the code (its
result was disregarded). So, the type ('a,unit) monad for it is far
better. We have changed the code and the paper to reflect that. There
is no longer `superfluous' unitL. Thank you for making this point.

> p23, what does "GenLA" do?
It is explained now, p30.

> p23, it would be nice to see the module signatures for the other
> features not discussed thus far
We added a new section, Sec 5.5, with the description of the other
features and their signatures.  Perhaps the code can be moved to an
appendix though?

> p24, missing "perform" in "init"
Actually, it is on the first line.

> p24, "module O = F.Output(F)" the functor "Output" takes a "OUTPUTDEP" not a
> "FEATURES", right?
Right, we explained that in more detail (see the paragraph before code
on p32).